---
title: "Conakry Cases, September 2014"
author: "CDC/OMS"
date: "Wednesday, October 22, 2014"
output:
  pdf_document:
    number_sections: yes
    toc: yes
fig_caption: yes
classoption: landscape
---

```{r, echo=FALSE, message=FALSE}
# linkage

library(stringr)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)
library(knitr)

opts_chunk$set( echo=FALSE, message=FALSE, comment="", warnings=FALSE, results='asis',
                cache=TRUE)
```

```{r read}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.

dir = 'fieldData/'  # folder, relative to this project, with spreadsheets

# it still may be possible to read csv..
# case_data =   read.csv( paste0(dir, file.name, '.csv'), header = TRUE, flush=TRUE )

# read CASES from file saved as excel
library(XLConnect)

# Case data
file.name = "guinea_cases_10_Nov"
excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
case_data =   readWorksheet(wb, file.name)

# CONTACTS

# Contacts (vhf)
file.name = "guinea_contacts_10_Nov"
excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
contact_data.vhf = readWorksheet(wb, file.name)

# Contacts ('online')
contact_data.online = contact_data.vhf %>%
  select(GlobalRecordId, ID, ThisContactIsAlsoCase, 
         Surname, OtherNames, Gender,
         Age, HeadHousehold, Village, District, SubCounty, 
         SourceCaseID, SourceCase, DateLastContact, 
         HCW, TotalSourceCases, DateOfLastFollowUp, FinalOutcome)
nrow(contact_data.online)
max(contact_data.online$DateLastContact, na.rm=T)

# Contacts (offline)
file.name = "Contacts-OMS-offline2"
excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
contact_data.offline = readWorksheet(wb, 'Contacts-OMS-offline') %>%
  select(GlobalRecordId, ID, ThisContactIsAlsoCase, 
         Surname, OtherNames, Gender,
         Age, HeadHousehold, Village, District, SubCounty, 
         SourceCaseID, SourceCase, DateLastContact, 
         HCW, TotalSourceCases, DateOfLastFollowUp, FinalOutcome)
nrow(contact_data.offline)
max(contact_data.offline$DateLastContact, na.rm=T)

# merge online and offline db?

# contact_data.merged = rbind(contact_data.offline, contact_data.online) 
new.offline = anti_join(contact_data.online, contact_data.offline, by=c('ID' = 'ID') )
nrow(new.offline)
# 1456 records in online file not in offline file

new.online = anti_join(contact_data.offline, contact_data.online, by=c('ID' = 'ID') )
nrow(new.online)
# 0 records in offline file not in online file

### use online!!!
contacts = contact_data.vhf 

# set report date as the most recent DateReport
last.report.date = max(case_data$DateReport, na.rm=TRUE)
current.week = week(last.report.date)

# create ifelse function that preserves date class.  see: http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no) structure(ifelse(cond, yes, no), class = class(yes))

```

```{r Define_case_data}

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Cas','Confirme',
                                          'Probable','Suspect', 'Exclude'))

# if else with dates can be troublesome.  see sol at 
# http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no){ class.y <- class(yes)
                                  X <- ifelse(cond,yes,no)
                                  class(X) <-class.y; return(X)}

last.report.date = max(case_data$DateReport, na.rm=T) # last day with report
current.week = week(last.report.date)  

cases = case_data %>% 
  select( GlobalRecordId, ID, ThisCaseIsAlsoContact,
          Surname, OtherNames,
          DateIsolationCurrent,
          StatusAsOfCurrentDate, EpiCaseDef , 
          DateReport, Age, AgeUnit, Gender, DateDeath, 
          StatusReport, VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, PlaceDeath, 
          HCWposition, HCWFacility, 
          DateOnsetLocalStart, DateOnset,
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EpiCaseDef==4) %>% # drop 'excluded'
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  inner_join(epiCaseDef) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          SCRes = toupper(SCRes),
          DistrictRes = toupper(DistrictRes)
          ) %>%
    filter( EpiCaseDef.label %in% c('Confirme','Probable', 'Suspect'))

# ensure that character encoding of imported data is 'UTF-8'.  Default on some windows machines is 'latin1'.  This will cause conflict when trying to join with the Region data.table (whose names are in UTF-8)
# cases$DistrictRes = iconv(cases$DistrictRes, 'latin1', 'UTF-8')
# or 
# Encoding(cases$VillageRes) = 'UTF-8'
# Encoding(cases$SCRes) = 'UTF-8'

```

```{r ckyCases}
load("regions.rda")

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

# focus on cases and contact in September (start by including those since mid August)
cky.cases = cases %>% 
  select( ID, Surname, OtherNames, dateOnset, DistrictRes, SCRes, VillageRes, Age) %>%
  filter(dateOnset>=ymd("2014-SEP-1") &
                               DistrictRes %in% "CONAKRY") %>% 
  arrange(dateOnset)

# group by districts
cky.cases %>%
  group_by( SCRes) %>%
  summarize( freq = n() )

# find cases with missing sc
cat('Table of Conakry cases with missing SCRes (sub-county)')
kable( 
  cky.cases %>% filter(is.na(SCRes)) 
  )

```

```{r contacts_clean}
cases.brief = case_data %>% select(ID, EpiCaseDef, Surname, OtherNames, Age, Gender, 
                                   HeadHouse, VillageRes, SCRes, DistrictRes) %>%
  mutate( SCRes = toupper(SCRes)) %>%
  rename(caseSurname = Surname,
         caseOtherNames = OtherNames,
         caseAge = Age,
         caseGender = Gender)

contacts= contacts %>% 
  inner_join(cases.brief, by = c('SourceCaseID' = 'ID')) %>% 
  arrange(DateLastContact) %>%
  mutate( idnum = substrRight(ID, 4),
          period = as.Date(floor_date(DateLastContact, "week")),
          SubCounty = toupper(SubCounty),
          Village = toupper(Village)) 


# clean district names with fuzz matching (agrepl)
district = 'RATOMA'
contacts[ agrepl(district, contacts$SubCounty, max =1), 'SubCounty' ] = district
district = 'MATOTO'
contacts[ agrepl(district, contacts$SubCounty, max =1), 'SubCounty' ] = district
district = 'MATAM'
contacts[ agrepl(district, contacts$SubCounty, max =1), 'SubCounty' ] = district
district = 'KALOUM'
contacts[ agrepl(district, contacts$SubCounty, max =1), 'SubCounty' ] = district
district = 'DIXINN'
contacts[ agrepl(district, contacts$SubCounty, max =1), 'SubCounty' ] = district

# special cases
exceptions = c('COLEAH', 'CONAKRY', 'DR FANTA BINTOU', 'M', 'SANGOYA', 'SONFONIA GARE')
contacts[ contacts$SubCounty %in% exceptions, 'SubCounty' ] = NA

```

\newpage

```{r epicurve_cases_table, fig.height=7, fig.width=9}
t = with(cky.cases, table(SCRes, DistrictRes))
 kable( t  )
```

NOTE: For `r sum(is.na(cky.cases$SCRes))` cases, DistrictRes is missing

```{r epicurve_cases, fig.height=7, fig.width=9}

cky.cases %>%
#   filter(
#              agrepl(c('hafia'), VillageRes, ignore.case=TRUE ) |
#               agrepl(c('cobaya'), VillageRes, ignore.case=TRUE ) |
#                agrepl(c('simanbossia'), VillageRes, ignore.case=TRUE ) |
#                agrepl(c('matoto'), VillageRes, ignore.case=TRUE )
#          ) %>%
  mutate( idnum = substrRight(ID, 4),
          period = as.Date(floor_date(dateOnset, "week")),
          SCRes = toupper(SCRes),
          VillageRes = toupper(VillageRes)) %>% 
  group_by( period, SCRes, idnum ) %>%
  summarize( freq = n() ) %>%  
  mutate( cumsum = cumsum(freq),
          midpoint = (cumsum(freq) - 0.5 * freq)) %>%
  ## plot
  ggplot( aes(x = period , y = freq) ) +
  geom_bar( stat = 'identity', aes(color = SCRes), fill='white') +
  geom_text( aes(y = midpoint, ymax=midpoint, label = idnum), size = 2, hjust=1) +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) +
  facet_grid(SCRes ~ .)

```

```{r epicurve_contacts, fig.height=7, fig.width=9}

# create table of cases with only a few variables to join with contacts. rename vars to prevent duplicate names

cky.contacts= contacts %>% 
  filter(DateLastContact>=ymd("2014-SEP-1") &
                               tolower(District) == "conakry") 

# group by districts
cky.contacts %>%
  group_by( SCRes) %>%
  summarize( freq = n() )

# group by period
cky.contacts %>%
  group_by( period) %>%
  summarize( freq = n() )

# 'Fix' data for contact  C-4223, with DateLastContact == 2014-11-30.  
# source case, gui-cky-14-3833 was not a case (discharged from donka after 3 neg tests) 
## Are there other contacts to non-cases?
# _ how many conatacts are missing source case id? none.  
# sum(is.na(cky.contacts$SourceCaseID))
# link contacts with source case and view EpiCaseDef

# cat('When linking contacts to their source case, some were not cases!')
# kable(
#   cky.contacts  %>% count(EpiCaseDef)
#   )

# filter out contacts of non cases
cky.contacts = cky.contacts %>% 
  filter( EpiCaseDef %in% c(1:3))

# cat('filtering out those whose source case were determined to be "non-case", there are',
#    nrow(cky.contacts) , 'contacts to', length(unique(cky.contacts$SourceCaseID)), 'cases.' )

cky.contacts %>% 
  group_by( period, SubCounty, idnum ) %>%
  summarize( freq = n() ) %>%  
  mutate( cumsum = cumsum(freq),
          midpoint = (cumsum(freq) - 0.5 * freq)) %>%
  ## plot
  ggplot( aes(x = period , y = freq) ) +
  geom_bar( stat = 'identity',  aes(color = SubCounty, fill=SubCounty)) +
#   geom_text( aes(y = midpoint, ymax=midpoint, label = idnum), size = 2, hjust=1) +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Ccontacts') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) +
  facet_grid(SubCounty ~ .)

```

```{r location_table}

kable(
  with(cky.contacts, table(SubCounty, DistrictRes))
  )

cat('NOTE: For', sum(is.na(cky.contacts$SubCounty)) ,'contacts, SubCounty is missing')

kable(
  with(cky.contacts, table(SubCounty, SCRes))
  )

cat('NOTE: For', sum(is.na(cky.contacts$SCRes)) ,'contacts, SCRes of source cases is missing')

cky.sc = unique(cky.contacts$SubCounty)
cat('Of the', nrow(cky.contacts), 'contacts in Conakry', 
    nrow(cky.contacts %>% filter(DistrictRes %in% 'Conakry')),
    'had source cases in Conakry,',
    nrow(cky.contacts %>% filter( DistrictRes %in% 'Coyah' )),
    'had source cases from Coyah, and source case '
    )

```

```{r source_cases, fig.height=7, fig.width=9}
cky.contacts %>% 
  group_by( period, SCRes, SourceCaseID ) %>%
  summarize( freq = n() ,
             sum = 1) %>%  
  mutate( cumsum = cumsum(sum),
          midpoint = (cumsum(sum) - 0.5 * sum),
          caseid = substrRight(SourceCaseID, 4)) %>%
  ## plot
  ggplot( aes(x = period, y = sum ) ) +
  geom_bar( stat = 'identity', color = 'black', fill='white') +
  geom_text( aes(y = midpoint, ymax=midpoint, 
                 label = paste( caseid , '(', freq ,')' ) ), size = 2, hjust=1) +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Contacts') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) +
  facet_grid(SCRes ~ .)
```

\newpage

```{r case_list}

cat('there are', nrow(cky.cases), 'cases after Sep 1 and ', 
    nrow(cky.contacts), ' contacts\n' )

# pick columns that need to be looked at to determine match
case.cols = c('ID', 'Surname','OtherNames', 'Age', 'DistrictRes', 'SCRes', 'VillageRes', 'dateOnset')
contact.cols = c('ThisContactIsAlsoCase', 'ID', 'Surname','OtherNames', 'Age', 'District', 'SubCounty', 'Village', 'DateLastContact')

kable(cky.cases[,case.cols])
```


# Potential Case-Contact Matches


```{r match2}
# find cky.cases in contacts:
# Exact string matching

# A) match surname to surname and othernames to othernames.  Find any where both are true
# match surname-surname

# expand list of contacts to begining of August
cky.contacts = contacts %>%
  filter(DateLastContact>=ymd("2014-AUG-1") &
                               tolower(District) == "conakry") 

contactMatch = function(.fun=grepl, ...){
  # match surname-surname
  sur.sur = sapply(cky.cases$Surname, .fun, cky.contacts$Surname, ignore.case=TRUE )
  
  # match othernames-othernames
  other.other = sapply(cky.cases$OtherNames, .fun, cky.contacts$OtherNames, ignore.case=TRUE )
  
  # matching surnames-othernames
  other.sur = sapply(cky.cases$OtherNames, .fun, cky.contacts$Surname, ignore.case=TRUE )
  
  # matching othernames-surnames
  sur.other = sapply(cky.cases$Surname, .fun, cky.contacts$OtherNames, ignore.case=TRUE )
  
  # both true when surname match with surname, or both true when surname match with OtherNames
  match = sur.sur * other.other + other.sur * sur.other
  
  # inspect matches
  return = list(
    case.match =  unique(which(match>=1, arr.ind=TRUE)[,2]),
    contact.match = unique(which(match>=1, arr.ind=TRUE)[,1]) ,
    match 
  )
}

x = contactMatch(.fun = agrepl, max=3)

# length(x[[1]]); length(x[[2]])
cat('there are', length(x[[1]]), 'cases with', 
    length(x[[2]]), 'matching contacts')

# match matrix
match = x[[3]]

which.match = which(match>=1, arr.ind=TRUE)
case.matches = unique(which.match[, 'col'])

# setup match fields
cky.cases$WasContact = 0L
cky.contacts$BecameCase = 0L

# select columns
case.cols = c('ID', 'Surname','OtherNames', 'Age', 'DistrictRes', 'SCRes', 'VillageRes', 'dateOnset')

contact.cols = c('ThisContactIsAlsoCase', 'ID', 'Surname','OtherNames', 'Age', 'District', 'SubCounty', 'Village', 'DateLastContact', 'SourceCase', 'SourceCaseID')

# print all combos
for (i in 1:length(x[[1]])){
  cat('\n\n **Match number',i, '**\n')
  cky.cases[ case.matches[i], 'WasContact'] = 1
  firstContactMatch = which.match[which(which.match[,'col']==case.matches[i]) , 'row'][1]
  cky.cases[ case.matches[i], 'ContactID'] = cky.contacts[ firstContactMatch ,
                                                                     'ID']
  cky.cases[ case.matches[i], 'SourceCaseID'] = cky.contacts[ firstContactMatch ,
                                                                     'SourceCaseID']
  print ( kable( cky.cases[ case.matches[i] , case.cols] ) )
  cat('contact:\n')
  cky.contacts[ which.match[which(which.match[,'col']==case.matches[i]) , 'row'], 'BecameCase'] = 1
  print( kable( cky.contacts[ which.match[which(which.match[,'col']==case.matches[i]) , 'row'], 
                              contact.cols]) )
}

save(cky.cases, cky.contacts, file='cky.cases.rda')

```

```{r secondary_cases}
john.ben.interval = new_interval(ymd("2014-SEP-1"), ymd("2014-OCT-30"))

t = cky.cases %>% 
  filter(dateOnset %within% john.ben.interval ) %>%
  count(SourceCaseID) %>%
  arrange(-n) %>%
  rename(Number_2ndary_Cases = n)

kable(t)

total.secondary = sum(t$Number_2ndary_Cases, na.rm=T)
number.sources = sum(!is.na(t$SourceCaseID))

t %>% filter(!is.na(t$SourceCaseID)) %>%
ggplot(aes(x=Number_2ndary_Cases )) +
  geom_bar() +
  scale_x_continuous('Number of secondary cases', 
                     breaks = seq(1, max(t$Number_2ndary_Cases), 2)
                     ) 
#   geom_annotate(x=7, y=10, type='text', text=paste('))
```


```{r epicurve_cases_were_contacts, fig.height=7, fig.width=9}
t = with(cky.cases, table(SCRes, WasContact))
kable( t  )

# cat('NOTE: For', sum(is.na(cky.cases$SCRes)) ,'cases, DistrictRes is missing')

sourceCases = cases %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( idnum = substrRight(ID, 4),
          period = as.Date(floor_date(dateOnset, "week")),
          SCRes = toupper(SCRes),
          VillageRes = toupper(VillageRes),
          WasContact = factor(WasContact, c(0,1), c('No','Yes'))
          ) %>% 
  group_by( period, SCRes, idnum, WasContact, sourceSCRes ) %>%
  summarize( freq = n() ) %>%  
  mutate( cumsum = cumsum(freq),
          midpoint = (cumsum(freq) - 0.5 * freq)) %>%
  ## plot
  ggplot( aes(x = period , y = freq) ) +
  geom_bar( stat = 'identity', aes(color = sourceSCRes, fill=sourceSCRes) ) +
  scale_fill_discrete('Case Was a Contact of Case in:') + # values = c('grey90', 'black')) +
  scale_color_discrete('Case Was a Contact of Case in:') +
  geom_text( aes(y = midpoint, ymax=midpoint, label = idnum), size = 2, hjust=1) +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) +
  facet_grid(SCRes ~ .)

```


# Donka registry

```{r donka}
library(XLConnect)
wb = loadWorkbook('fieldData/2014_10_25_Registre Donka Version 4-5.2.xlsx')
donka = readWorksheet(wb, 'registre') %>% 
  filter( Statut == 'Confirmé') 

donka.sept = donka %>% filter( Date.Admis %within% 
                                 interval(ymd('2014-Sep-01'), ymd('2014-OCT-31')) 
                               ) %>%
          select( Num, Nom..Prénom..âge..genre, Age, Genre, Commentaires, Date.Symptômes, Sorties, Durée.de.séjour, quartiers.a.precise, quartier, Origine) %>%
  arrange( Date.Symptômes )


donka.sept.locations = donka %>% filter( Date.Admis %within% 
                                 interval(ymd('2014-Sep-01'), ymd('2014-OCT-31')) 
                               ) %>%
          select( Num, quartiers.a.precise, quartier, Origine) %>%
  arrange( Num )


admitted = nrow(donka.sept)
died = nrow(donka.sept %>% filter(Sorties == 'Décédé'))

admitted.all = nrow(donka)
died.all = nrow(donka %>% filter(Sorties == 'Décédé'))
                                  
kable(donka.sept)
kable(donka.sept.locations)
```

In September, there were `r admitted` cases admitted to Donka hospital.  Of these, `r died` (`r percent(died/admitted)` ) survived. Since the beginning of the outbreak, there have been `r admitted.all` cases admitted
and `r died` (`r percent(died.all/admitted.all)` ) survived.