---
title: "dashboard"
author: "CDC/OMS"
date: "Monday, November 24, 2014"
output: beamer_presentation
---
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warning=FALSE, results='asis')

library(XLConnect, quietly=TRUE)
library(lubridate)
library(reshape)  # use for rename() function

suppressPackageStartupMessages(library(ggplot2))
library(scales)
library(grid)
library(RColorBrewer)

library(tidyr) # install from github (CRAN version not working...)
library(dplyr, quietly=TRUE) # order is important...

```

# WHO indicators

```{r read_who}
# read data from excel 

dir = ''
file.name = 'WHO indicators'
# read from file saved as excel
excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)

labels =   readWorksheet(wb, 'labels')
weeks =  readWorksheet(wb, 'weeks')
guinea = readWorksheet(wb, 'guinea')
SL = readWorksheet(wb, 'SL')
liberia = readWorksheet(wb, 'liberia')

```

## WHO Indicators, Guinea

```{r guinea, warning=FALSE}
currentWeekColumn = ncol(guinea)

guinea = guinea %>% 
  inner_join(labels) %>% 
  mutate(
    change = guinea[, currentWeekColumn] - guinea[, currentWeekColumn - 1],
    value = ifelse(type %in% 'percent', guinea[,currentWeekColumn], NA) 
    )

guinea$label = factor(guinea$label, levels = rev(guinea[guinea$order, 'label']) )
                                            
bars = guinea %>%
  ggplot(aes(x=label, y=100, order = order)) +
  ggtitle('Most Recent Week') +
  ylab('') +
  xlab('') +
  geom_bar(fill = 'grey', alpha = .9, stat = 'identity') +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity', aes(y = value) ) +
  geom_text(color = 'white', face = 'bold', size = 12,
            aes(y = 50,label = guinea[,currentWeekColumn] ) ) +
  coord_flip() +
  theme(axis.text= element_text(size=24, color = 'black'),
        axis.title = element_text(size=32, face='bold', color = 'black'))

guinea$label = factor(guinea$label, levels = guinea[guinea$order, 'label'])

sparklines = guinea %>% select(-value) %>%
  rename(var = variable) %>%
  gather(label, week, week46:week48)   %>% 
  mutate(value = ifelse(is.na(value), 0, value) ) %>%
  ggplot( aes( x= variable, y = value, group = label, order = order) ) +
  ggtitle('Change') +
  geom_line( color = 'blue') +
  geom_point(color = 'blue') +
#   annotate('text', aes(label = change)) +
  facet_grid(label ~ .,  scales = "free") +
  theme_bw() +
  xlab('') + ylab('') +
  theme( 
        axis.text= element_blank(),
         axis.line =         element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         axis.ticks = element_blank(),
         strip.background = element_blank(),
         panel.background = element_blank(),
         strip.background = element_blank(), 
         strip.text = element_blank(),
         plot.background =   element_rect(colour = NA),
         panel.border = element_blank()) 


# library(gridExtra)
# grid.arrange( bars, sparklines, ncol=2)

file.remove("dashGIN.png")
png(file="dashGIN.png", width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
grid.newpage()
  pushViewport(viewport(layout = grid.layout(1, 20)))
  vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
  print(bars, vp = vplayout(1, 1:17))
  print(sparklines, vp = vplayout(1, 18:20))
dev.off()
```

## WHO Indicators, Sierra Leone

```{r SL, warning=FALSE}
currentWeekColumn = ncol(SL)

SL = SL %>% 
  inner_join(labels) %>% 
  mutate(
    change = SL[, currentWeekColumn] - SL[, currentWeekColumn - 1],
    value = ifelse(type %in% 'percent', SL[,currentWeekColumn], NA) 
    )

SL$label = factor(SL$label, levels = rev(SL[SL$order, 'label']) )
                                            
bars = SL %>%
  ggplot(aes(x=label, y=100, order = order)) +
  ggtitle('Most Recent Week') +
  ylab('') +
  xlab('') +
  geom_bar(fill = 'grey', alpha = .9, stat = 'identity') +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity', aes(y = value) ) +
  geom_text(color = 'white', face = 'bold', size = 12, 
            aes(y = 50,label = SL[,currentWeekColumn] ) ) +
  coord_flip() +
  theme(axis.text= element_text(size=24, color = 'black'),
        axis.title = element_text(size=32, face='bold', color = 'black'))

SL$label = factor(SL$label, levels = SL[SL$order, 'label'])

sparklines = SL %>% select(-value) %>%
  rename(var = variable) %>%
  gather(label, week, week46:week48)   %>% 
  mutate(value = ifelse(is.na(value), 0, value) ) %>%
  ggplot( aes( x= variable, y = value, group = label, order = order) ) +
  ggtitle('Change') +
  geom_line( color = 'blue') +
  geom_point(color = 'blue') +
  facet_grid(label ~ .,  scales = "free") +
  theme_bw() +
  xlab('') + ylab('') +
  theme( 
        axis.text= element_blank(),
         axis.line =         element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         axis.ticks = element_blank(),
         strip.background = element_blank(),
         panel.background = element_blank(),
         strip.background = element_blank(), 
         strip.text = element_blank(),
         plot.background =   element_rect(colour = NA),
         panel.border = element_blank()) 


# library(gridExtra)
# grid.arrange( bars, sparklines, ncol=2)

file.remove("dashSL.png")
png(file="dashSL.png", width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
grid.newpage()
  pushViewport(viewport(layout = grid.layout(1, 20)))
  vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
  print(bars, vp = vplayout(1, 1:17))
  print(sparklines, vp = vplayout(1, 18:20))
dev.off()
```

## WHO Indicators, Liberia

```{r liberia, warning=FALSE}
currentWeekColumn = ncol(liberia)

liberia = liberia %>% 
  inner_join(labels) %>% 
  mutate(
    change = liberia[, currentWeekColumn] - liberia[, currentWeekColumn - 1],
    value = ifelse(type %in% 'percent', liberia[,currentWeekColumn], NA) 
    )

liberia$label = factor(liberia$label, levels = rev(liberia[liberia$order, 'label']) )
                                            
bars = liberia %>%
  ggplot(aes(x=label, y=100, order = order)) +
  ggtitle('Most Recent Week') +
  ylab('') +
  xlab('') +
  geom_bar(fill = 'grey', alpha = .9, stat = 'identity') +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity', aes(y = value) ) +
  geom_text(color = 'white', face = 'bold', size = 12, 
            aes(y = 50,label = liberia[,currentWeekColumn] ) ) +
  coord_flip() +
  theme(axis.text= element_text(size=24, color = 'black'),
        axis.title = element_text(size=32, face='bold', color = 'black'))

liberia$label = factor(liberia$label, levels = liberia[liberia$order, 'label'])

sparklines = liberia %>% select(-value) %>%
  rename(var = variable) %>%
  gather(label, week, week46:week48)   %>% 
  mutate(value = ifelse(is.na(value), 0, value) ) %>%
  ggplot( aes( x= variable, y = value, group = label, order = order) ) +
  ggtitle('Change') +
  geom_line( color = 'blue') +
  geom_point(color = 'blue') +
  facet_grid(label ~ .,  scales = "free") +
  theme_bw() +
  xlab('') + ylab('') +
  theme( 
        axis.text= element_blank(),
         axis.line =         element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         axis.ticks = element_blank(),
         strip.background = element_blank(),
         panel.background = element_blank(),
         strip.background = element_blank(), 
         strip.text = element_blank(),
         plot.background =   element_rect(colour = NA),
         panel.border = element_blank()) 


# library(gridExtra)
# grid.arrange( bars, sparklines, ncol=2)

file.remove("dashLIB.png")
png(file="dashLIB.png", width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
grid.newpage()
  pushViewport(viewport(layout = grid.layout(1, 20)))
  vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
  print(bars, vp = vplayout(1, 1:17))
  print(sparklines, vp = vplayout(1, 18:20))
dev.off()
```

# VHF

```{r vhf_data}

# Import data from combined sas data file:
# 1. export sas file to csv
# 2. import...
# 3. filter data to suspect, probable, and confirmed
# 4. adjust place names

dir = 'fieldData/'
file = paste0(dir, "case_guinea_sl.csv" ) 

options(stringsAsFactors=FALSE)
text = readLines(file, skipNul=TRUE, encoding="UTF-8", n=-1L )
case_guinea_sl = read.csv(textConnection(text), header = TRUE, stringsAsFactors = FALSE)

# convert character data to dates
case_guinea_sl = case_guinea_sl %>% 
  mutate(
          DateIsolationCurrent = dmy(DateIsolationCurrent) ,
          DateHospitalCurrentAdmit = dmy(DateHospitalCurrentAdmit),
          DateReport = dmy(DateReport), 
          DateOnset = dmy(DateOnset)
          )

# set timezone attributes
attr(case_guinea_sl$DateOnset, "tzone") <- "UTC"
attr(case_guinea_sl$DateReport, "tzone") <- "UTC"
attr(case_guinea_sl$DateHospitalCurrentAdmit, "tzone") <- "UTC"
attr(case_guinea_sl$DateIsolationCurrent, "tzone") <- "UTC"


EpiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Case','Confirmed',
                                          'Probable','Suspect', 'Excluded'))

last.report.date = max(case_guinea_sl$DateReport, na.rm=T)  # last day with report
current.week = week( last.report.date)

cases.guinea = case_guinea_sl %>% 
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  inner_join(EpiCaseDef) %>%  
  mutate(
          dateOnset = pmin( DateReport, DateOnset , na.rm = T),
          # use onset date if available; DateReport otherwise
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirmed','Probable', 'Suspect'))

# clean districtRes

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Gueckedou', 'DistrictRes'] = 'Guéckédou'
  
  case_guinea_sl[case_guinea_sl$DistrictRes %in% "N'Zerekore", 'DistrictRes'] = 'Nzérékoré'
  
  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Telimélé', 'DistrictRes'] = 'Télimélé'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Kerouane', 'DistrictRes'] = 'Kérouané'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Boke', 'DistrictRes'] = 'Boké'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Labe', 'DistrictRes'] = 'Labé'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Lelouma', 'DistrictRes'] = 'Lélouma'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Tougue', 'DistrictRes'] = 'Tougué'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Forecariah', 'DistrictRes'] = 'Forécariah'

  case_guinea_sl[case_guinea_sl$DistrictRes %in% 'Dubreka', 'DistrictRes'] = 'Dubréka'


# Create separate country datasets
case_guinea = case_guinea_sl %>% filter(Country %in% 'Guinea') %>% as.data.frame()
case_sl = case_guinea_sl %>% filter(Country %in% 'Sierra Leone') %>% as.data.frame()

save(case_guinea_sl, case_guinea,  case_sl, file='case_guinea_sl.rda')

```

# Proportion hospitalized within 3 days of onset

## Proportion hospitalized within 3 days of onset

```{r days_iso_proportion}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

load('case_guinea_sl.rda')
data = case_guinea_sl %>% 
  filter( DateOnset  >= ymd('2014-JUL-1')) %>%
  select( Country,
          DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( Country, 
            period = as.Date(floor_date(DateOnset, "week"))) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             )
             
 
g = ggplot(data , aes( period, within3daysToHosp)) +
    scale_x_date('\nWeek beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset\n', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text(color = 'white', size = 8,
            aes(label = n ), vjust = 2)  +
  theme(axis.text= element_text(angle = 90, color = 'black', size=24) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=32, face="bold")
        ) +
  facet_grid( Country ~ . )
g

file.remove('dash.perchosp.png')
png(file='dash.perchosp.png', width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
g
dev.off()
```

## Proportion hospitalized within 3 days of onset, coded

```{r days_iso_proportion_color}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

load('case_guinea_sl.rda')
data = case_guinea_sl %>% 
  filter( DateOnset  >= ymd('2014-JUL-1')) %>%
  select( Country,
          DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( Country, 
            period = as.Date(floor_date(DateOnset, "week"))) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             ) %>%
  mutate( lever = ifelse( within3daysToHosp > .7, 'green',
                          ifelse( within3daysToHosp > .5, 'yellow', 'red'))
          )
             
 
g = ggplot(data , aes( period, within3daysToHosp)) +
    scale_x_date('\nWeek beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset\n', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(alpha = .9, stat = 'identity', aes(fill = lever)) +
  scale_fill_manual(values = c('green' = 'green', 'yellow'='yellow', 'red' = 'red'), 
                    guide = FALSE ) +
  geom_text(color = 'white', size = 8,
            aes(label = n ), vjust = 2)  +
  theme(axis.text= element_text(angle = 90, color = 'black', size=24) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=32, face="bold")
        ) +
  facet_grid( Country ~ . )
g

file.remove('dash.perchosp.coded.png')
png(file='dash.perchosp.coded.png', width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
g
dev.off()
```

## Proportion hospitalized within 3 days of onset, Guinea, by district, coded

### by district

```{r days_iso_proportionfacet}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

load('case_guinea_sl.rda')
data = case_guinea %>% 
  filter( DateOnset  >= ymd('2014-JUL-1')) %>%
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( Region, 
            period = as.Date(floor_date(DateOnset, "week"))) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             ) %>%
  mutate( lever = ifelse( within3daysToHosp > .7, 'green',
                          ifelse( within3daysToHosp > .5, 'yellow', 'red'))
          ) %>%
  filter( !is.na(Region) )
             
 
g = ggplot(data , aes( period, within3daysToHosp)) +
    scale_x_date('\nWeek beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset\n', 
                       labels = percent, limits = c(0,1), breaks = .5) +
    geom_bar(aes( fill = lever) , stat = 'identity') +
    scale_fill_manual(values = c('green' = 'green', 'yellow'='yellow', 'red' = 'red'), 
                    guide = FALSE ) +
    geom_text(color = 'white', size = 4,
            aes(label = n ), vjust = 1)  +
    theme(axis.text= element_text(angle = 90, color = 'black', size=20) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=20, face="bold")
        ) +
  facet_grid( Region ~ .)
g

file.remove('dash.perchosp.guinea.facet.png')
png(file='dash.perchosp.guinea.facet.png', width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
g
dev.off()
```

## Detail of Days to Hospitalization, Guinea

```{r days_iso}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('case_guinea_sl.rda')
data = case_guinea %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) 
 
g = ggplot(data , aes( month(DateOnset), daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Hospitalization', breaks = seq(0, 28, 2), limits =c(0,14)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("grey",  "green", "red")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  +
    theme(axis.text= element_text(angle = 90, color = 'black', size=20) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=20, face="bold")
        ) 
g

file.remove('dash.perchosp.gin.detail.png')
png(file='dash.perchosp.gin.detail.png', width = 22, height = 10, units = 'in', res = 300)
g
dev.off()
```

## Detail of Days to Hospitalization, SL

```{r days_iso_sl}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.sl.rda')
data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) 
 
g = ggplot(data , aes( month(DateOnset), daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Hospitalization', breaks = seq(0, 28, 2), limits =c(0,14)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("grey",  "green", "red")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  
# + 
#   facet_grid( Region ~ .)
g

file.remove('dash.perchosp.sl.detail.png')
png(file='dash.perchosp.sl.detail.png', width = 22, height = 10, units = 'in', res = 300)
g
dev.off()
```

# Cases that were Contacts

## Cases that were Contacts, Sep-Oct, Conakry

```{r cases_found_among_contact, fig.height=6, fig.width=10, eval=FALSE}

load('cky.cases.rda')

# percent cases that were contacts
data = cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( 
          SCRes = toupper(SCRes),
          period = as.Date(floor_date(dateOnset, "week"))
          ) %>% 
  group_by( period ) %>%
  summarize( n = n(),
             contact = sum(WasContact, na.rm=T),
             caseContacts = contact/n
             ) %>%
    mutate( lever = ifelse( caseContacts > .7, 'green',
                          ifelse( caseContacts > .5, 'yellow', 'red'))
          )

g = ggplot(data , aes( period, caseContacts)) +
    scale_x_date('\nWeek beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
    scale_y_continuous('% Cases that were Contacts\n', 
                       labels = percent, limits = c(0,1)) +
    geom_bar(aes( fill = lever) , alpha = .9, stat = 'identity') +
      scale_fill_manual(values = c('green' = 'green', 'yellow'='yellow', 'red' = 'red'), 
                    guide = FALSE ) +
#   geom_text(color = 'white', size = 8,
#             aes(label = n ), vjust = 1)  +
    theme(axis.text= element_text(angle = 90, color = 'black', size=24) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=32, face="bold")
        ) 
g

file.remove('dash.perccontact.gin.png')
png(file='dash.perccontact.gin.png', width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
g
dev.off()
```

## Cases that were Contacts, Sep-Oct, Conakry, by sub-district

```{r epicurve_cases_were_contacts, fig.height=6, fig.width=10}

load('cky.cases.rda')

sourceCases = cases.guinea %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
# percent cases that were contacts
data = cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( 
          period = as.Date(floor_date(dateOnset, "week")),
          SCRes = toupper(SCRes)
          ) %>% 
  group_by( period, SCRes ) %>%
  summarize( n = n(),
             contact = sum(WasContact, na.rm=T),
             caseContacts = contact/n
             ) %>% 
  mutate( lever = ifelse( caseContacts > .7, 'green',
                          ifelse( caseContacts > .5, 'yellow', 'red'))
          ) %>%
  filter(!is.na(SCRes))

g = ggplot(data , aes( period, caseContacts)) +
    scale_x_date('\nWeek beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
    scale_y_continuous('% Cases that were Contacts\n', 
                       labels = percent, limits = c(0,1), breaks = .5) +
    geom_bar(aes( fill = lever) ,alpha = .9, stat = 'identity') +
      scale_fill_manual(values = c('green' = 'green', 'yellow'='yellow', 'red' = 'red'), 
                    guide = FALSE ) +
    geom_text( color = 'white', aes(y = caseContacts/2, label = percent(caseContacts) )) +
    theme(axis.text= element_text(angle = 90, color = 'black', size=24) ,
        axis.title= element_text(face = 'bold', color = 'black', size=32),
        strip.text = element_text(size=28, face="bold")
        ) +
  facet_grid(SCRes ~ .)
g

file.remove('dash.perccontact.gin.district.png')
png(file='dash.perccontact.gin.district.png', width = 24, height = 11, units = 'in', 
    res = 72, bg = "transparent")
g
dev.off()
```
