---
title: "dashboard"
author: "CDC/OMS"
date: "Monday, November 24, 2014"
output: beamer_presentation
---
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warning=FALSE, results='asis')

library(XLConnect, quietly=TRUE)
library(lubridate)
library(reshape)  # use for rename() function

suppressPackageStartupMessages(library(ggplot2))
library(scales)
library(grid)
library(RColorBrewer)

library(tidyr) # install from github (CRAN version not working...)
library(dplyr, quietly=TRUE) # order is important...

```


```{r read}
# read data from excel 

dir = ''
file.name = 'WHO indicators'
# read from file saved as excel
excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)

labels =   readWorksheet(wb, 'labels')
weeks =  readWorksheet(wb, 'weeks')
guinea = readWorksheet(wb, 'guinea')
SL = readWorksheet(wb, 'SL')
liberia = readWorksheet(wb, 'liberia')

```

## WHO Indicators, Guinea

```{r guinea, warning=FALSE}
currentWeekColumn = ncol(guinea)

guinea = guinea %>% 
  inner_join(labels) %>% 
  mutate(
    change = guinea[, currentWeekColumn] - guinea[, currentWeekColumn - 1],
    value = ifelse(type %in% 'percent', guinea[,currentWeekColumn], NA) 
    )

guinea$label = factor(guinea$label, levels = rev(guinea[guinea$order, 'label']) )
                                            
bars = guinea %>%
  ggplot(aes(x=label, y=100, order = order)) +
  ggtitle('Most Recent Week') +
  ylab('') +
  xlab('') +
  geom_bar(fill = 'grey', alpha = .9, stat = 'identity') +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity', aes(y = value) ) +
  geom_text(color = 'white', face = 'bold',
            aes(y = 50,label = guinea[,currentWeekColumn] ) ) +
  coord_flip() +
  theme(axis.text= element_text(size=12, face='bold', color = 'black'))

guinea$label = factor(guinea$label, levels = guinea[guinea$order, 'label'])

sparklines = guinea %>% select(-value) %>%
  rename(var = variable) %>%
  gather(label, week, week46:week47)   %>% 
  mutate(value = ifelse(is.na(value), 0, value) ) %>%
  ggplot( aes( x= variable, y = value, group = label, order = order) ) +
  ggtitle('Change') +
  geom_line( color = 'blue') +
  geom_point(color = 'blue') +
  facet_grid(label ~ .,  scales = "free") +
  theme_bw() +
  xlab('') + ylab('') +
  theme( 
        axis.text= element_blank(),
         axis.line =         element_blank(),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         axis.ticks = element_blank(),
         strip.background = element_blank(),
         panel.background = element_blank(),
         strip.background = element_blank(), 
         strip.text = element_blank(),
         plot.background =   element_rect(colour = NA),
         panel.border = element_blank()) 


# library(gridExtra)
# grid.arrange( bars, sparklines, ncol=2)

grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 10)))
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
print(bars, vp = vplayout(1, 1:9))
print(sparklines, vp = vplayout(1, 10))

```


## Proportion hospitalized within 3 days of onset, Guinea

```{r days_iso_proportion}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( Month = month(DateOnset)) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             )
             
 
g = ggplot(data , aes( Month, within3daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text(y = 50, color = 'white', aes(y = within3daysToHosp/2, label = percent(within3daysToHosp) )) 
g
```


## Proportion hospitalized within 3 days of onset, Guinea

### by district

```{r days_iso_proportionfacet}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( DistrictRes, Month = month(DateOnset)) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             ) %>% 
  filter(n>20)
             
 
g = ggplot(data , aes( Month, within3daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text(y = 50, color = 'white', aes(y = within3daysToHosp/2, label = percent(within3daysToHosp) )) +
  facet_grid(DistrictRes ~ .)
g
```

## Detail of Days to Hospitalization, Guinea

```{r days_iso}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) 
 
g = ggplot(data , aes( month(DateOnset), daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Hospitalization', breaks = seq(0, 28, 2), limits =c(0,14)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("grey",  "green", "red")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  
# + 
#   facet_grid( Region ~ .)
g

```

## Proportion hospitalized within 3 days of onset, SL

```{r days_iso_proportion_sl}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.sl.rda')
data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) %>%
  group_by( Month = month(DateOnset)) %>%
  summarize( n = n(),
             within3daysToHosp = sum(ifelse(daysToHosp>3, 0, 1), na.rm=T) / n 
             )
             
 
g = ggplot(data , aes( Month, within3daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text(y = 50, color = 'white', aes(y = within3daysToHosp/2, label = percent(within3daysToHosp) )) 
g
```


## Detail of Days to Hospitalization, SL

```{r days_iso_sl}

load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.sl.rda')
data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, DateHospitalCurrentAdmit, 
          StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToHosp = as.numeric( as.Date(DateHospitalCurrentAdmit) - as.Date(DateOnset) ) ) 
 
g = ggplot(data , aes( month(DateOnset), daysToHosp)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Hospitalization', breaks = seq(0, 28, 2), limits =c(0,14)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("grey",  "green", "red")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  
# + 
#   facet_grid( Region ~ .)
g

```


## Cases that were Contacts, Sep-Oct, Conakry

```{r cases_found_among_contact, fig.height=4, fig.width=10}

load('cky.cases.rda')

sourceCases = cases.guinea %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
# percent cases that were contacts
data = cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( 
          Month = month(dateOnset),
          SCRes = toupper(SCRes)
          ) %>% 
  group_by( Month ) %>%
  summarize( n = n(),
             contact = sum(WasContact, na.rm=T),
             caseContacts = contact/n
             ) 

g = ggplot(data , aes( Month, caseContacts)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text( color = 'white', aes(y = caseContacts/2, label = percent(caseContacts) )) 
g

```


## Cases that were Contacts, Sep-Oct, Conakry

### by sub-district

```{r epicurve_cases_were_contacts, fig.height=6, fig.width=10}

load('cky.cases.rda')

sourceCases = cases.guinea %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
# percent cases that were contacts
data = cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( 
          Month = month(dateOnset),
          SCRes = toupper(SCRes)
          ) %>% 
  group_by( Month, SCRes ) %>%
  summarize( n = n(),
             contact = sum(WasContact, na.rm=T),
             caseContacts = contact/n
             ) 

g = ggplot(data , aes( Month, caseContacts)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('% Hospitalized with 3 days of illness onset', 
                       labels = percent, limits = c(0,1)) +
  geom_bar(fill = 'blue', alpha = .9, stat = 'identity') +
  geom_text( color = 'white', aes(y = caseContacts/2, label = percent(caseContacts) )) +
  facet_grid(SCRes ~ .)
g
```
