---
title: "Under 5 Case Definition Variables"
author: "Epi Task Force"
date: "Thursday, December 04, 2014"
output: pdf_document
---

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
library(mi)
library(knitr)
library(scales)
library(dplyr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", warning=FALSE, results='asis')

load('case_guinea_sl.rda')

EpiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Case','Confirmed',
                                          'Probable','Suspect', 'Excluded'))

cases = case_guinea_sl %>% 
  select(AgeYears, Fever, Vomiting, Diarrhea, AbdPain, Unexplainedbleeding, Contact, EpiCaseDef) %>%
  filter( AgeYears<5) %>% 
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  left_join(EpiCaseDef)  %>%
  rename( Case = EpiCaseDef.label) %>%
  select(-EpiCaseDef)


under5 = nrow(cases)
```

Ebola case status and symptom records are available from VHF records of `r comma(under5)` children 0-4 years of age in Guinea and Sierra Leone. Hhowever, much of the data elements have missing values.  Available data for select variables is blue; missing data is red.


```{r}

par(mai=c(1,2,0,0.1))

cases %>% 
  missing.pattern.plot( clustered=FALSE,
                      xlab= paste0("observations (n=", comma(nrow(cases)), ")" ),
                      main = ''
                      )

```

Among those with available data, this table shows the percentages of children with known Ebola contact and other symptoms.

```{r }
library(data.table)
library(dplyr)

cases %>% 
  filter( Contact %in% c(1,2)) %>% 
#   mutate( Contact = factor(Contact, labels = c('Y', 'N'))) %>%
  as.data.table() %>%
  group_by( Case) %>%
  summarize( 
    n = n(),
    Contact = percent( sum(Contact == 1, na.rm = T)  / sum(Contact %in% c(1,2), na.rm = T) ) ,
    Fever = percent( sum(Fever == 1, na.rm = T)  / sum(Fever %in% c(1,2), na.rm = T) ) , 
    Vomiting = percent( sum(Vomiting == 1, na.rm = T)  / sum(Vomiting %in% c(1,2), na.rm = T) ) , 
    Diarrhea = percent( sum(Diarrhea == 1, na.rm = T)  / sum(Diarrhea %in% c(1,2), na.rm = T) ) , 
    AbdPain = percent( sum(AbdPain == 1, na.rm = T)  / sum(AbdPain %in% c(1,2), na.rm = T) ) , 
    Unexplainedbleeding = percent( sum(Unexplainedbleeding == 1, na.rm = T)  / sum(Unexplainedbleeding %in% c(1,2), na.rm = T) ) 
    ) %>%
  kable()


```

*n: indicates number of records with contact listed as yes or no only.  
