---
title: "Guinee"
author: "CDC/OMS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes
  word_document:
  - toc: yes
  
header-includes:
- \usepackage[section]{placeins}
- \usepackage{setspace}
- \usepackage{relsize}
---

```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }
options(tz="GMT+0")
Sys.setenv(TZ='GMT+0')

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warnings=FALSE, results='asis')

library(XLConnect, quietly=TRUE)
library(lubridate)
library(reshape)  # use for rename() function

library(rgeos)
library(sp)

library(ggplot2)
library(scales)
library(grid)
library(RColorBrewer)

# library(devtools)
# devtools::install_github("hadley/tidyr")
library(tidyr) # install from github (CRAN version not working...)
library(dplyr, quietly=TRUE) # order is important...

```


# Data 

- Data loaded from export of VHF database.^[export cases as .csv, then open and save file as excel]

```{r read}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.

file.name = "guinea_cases_19_Oct"

dir = 'fieldData\\'  # folder, relative to this project, with spreadsheets

excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
case_data =   readWorksheet(wb, file.name)
  
# set report date as the most recent DateReport
last.report.date = max(case_data$DateReport, na.rm=TRUE)
current.week = week(last.report.date)

# create ifelse function that preserves date class.  see: http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no) structure(ifelse(cond, yes, no), class = class(yes))

```

```{r Define_case_data}

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Cas','Confirme',
                                          'Probable','Suspect', 'Exclude'))

cases = case_data %>% 
  select( GlobalRecordId, ID, ThisCaseIsAlsoContact,
          Surname, OtherNames,
          StatusAsOfCurrentDate, EpiCaseDef , 
          DateReport, Age, AgeUnit, Gender, DateDeath, 
          StatusReport, VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, 
          HCWposition, HCWFacility, 
          DateOnsetLocalStart, DateOnset,
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EpiCaseDef==4) %>% # drop 'excluded'
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  inner_join(epiCaseDef) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirme','Probable', 'Suspect'))

# ensure that character encoding of imported data is 'UTF-8'.  Default on some windows machines is 'latin1'.  This will cause conflict when trying to join with the Region data.table (whose names are in UTF-8)
# cases$DistrictRes = iconv(cases$DistrictRes, 'latin1', 'UTF-8')
# or 
# Encoding(cases$VillageRes) = 'UTF-8'
# Encoding(cases$SCRes) = 'UTF-8'

```

```{r Clean_Locations}
  cases[cases$DistrictRes %in% 'Gueckedou', 'DistrictRes'] = 'Guéckédou'
  
  cases[cases$DistrictRes %in% "N'Zerekore", 'DistrictRes'] = 'Nzérékoré'
  
  cases[cases$DistrictRes %in% 'Telimélé', 'DistrictRes'] = 'Télimélé'

    cases[cases$DistrictRes %in% 'Kerouane', 'DistrictRes'] = 'Kérouané'

    cases[cases$DistrictRes %in% 'Boke', 'DistrictRes'] = 'Boké'

    cases[cases$DistrictRes %in% 'Labe', 'DistrictRes'] = 'Labé'

    cases[cases$DistrictRes %in% 'Lelouma', 'DistrictRes'] = 'Lélouma'

    cases[cases$DistrictRes %in% 'Tougue', 'DistrictRes'] = 'Tougué'

    cases[cases$DistrictRes %in% 'Forecariah', 'DistrictRes'] = 'Forécariah'

    cases[cases$DistrictRes %in% 'Dubreka', 'DistrictRes'] = 'Dubréka'

# save cases data
save(cases, file='cases.rda')

```

```{r missing_date, results='asis'}
#  list records missing period
missingPeriod = cases %>% filter(is.na(dateOnset)) %>% select(ID)
```

- The most recent case included in this report had symptom onset on 
**`r format(max(cases$DateReport, na.rm=TRUE), "%d-%B-%Y")`**.

- There are `r nrow(missingPeriod)` records with missing dates:   
`r cat(missingPeriod$ID, sep=', ')`

# Epicurve by Month

```{r epicurve_monthly, fig.cap='Monthly Ebola Cases, Guinee'}

# epicurve by month (2014 only) 

cases %>% 
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "month")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('month', labels = date_format("%b-%dd"),
               breaks = date_breaks("month")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve0.png')
  

```

# Epicurves By Week (by report date and onset date)

```{r epicurves_weekly_DateReported, fig.height=3.5, fig.cap='Weekly Ebola Cases, Guinee (by date reported)'}

cases %>% 
  filter( year(DateReport) == 2014 ) %>%
  arrange(DateReport) %>%
  mutate( interval = week(DateReport),
          period = as.Date(floor_date(DateReport, "week")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve1.png')
```

```{r epicurves_weekly_onsetDate, fig.height=3.5, fig.cap='Weekly Ebola Cases, Guinee (by onset date)'}
cases %>% 
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "week")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve2.png')

```

# Case-Status by Week

```{r tab_status_byOnsetWeek}

t.period_wk = cases %>%    # tabulate margin totals by status
  filter(!is.na(period_wk)) %>%
  group_by(Status) %>% 
  summarise(Total = n())

t = cases %>%  # tabulate totals by status and period_wk
  filter(!is.na(period_wk)) %>%
  group_by(Status, period_wk) %>% 
  summarise(Total = n()) %>% 
  as.data.frame() %>%  # spread works more reliably on data.frame
#   mutate( Percent = percent(Total/sum(Total, na.rm=TRUE)) ) %>%
  spread(period_wk, Total)  %>%
  inner_join(t.period_wk) 

# tabulate margin totals over status
TOTALS = cbind(Status = 'TOTAL',
                      data.frame( t(sapply(t[,2:ncol(t)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(t)  

t = rbind(t, TOTALS)  # join TOTALS to t

t[is.na(t)] <- 0  # convert NA to 0

# Calculate number of days reported in current week
daysThisWeek = cases %>% 
  select(period_wk, dateOnset) %>% 
  filter(
    if(sum(period_wk %in% 'Week')>0){period_wk=='Week'}else{period=='Week_1'}
       ) %>% 
  summarise( 
    days = max(dateOnset)- min(dateOnset) + 1 
    )

cat('The current week has', daysThisWeek$days, 'days\n')

kable(t)  # print table

```

# Case-Status by 7-day period, begining with last reported case

```{r tab_status_7days}

t.period = cases %>%    # tabulate margin totals by status
  filter(!is.na(period)) %>%
  group_by(Status) %>% 
  summarise(Total = n())

t = cases %>%  # tabulate totals by status and period
  filter(!is.na(period)) %>%
  group_by(Status, period) %>% 
  summarise(Total = n()) %>% 
  as.data.frame() %>%  # spread works more reliably on data.frame
#   mutate( Percent = percent(Total/sum(Total, na.rm=TRUE)) ) %>%
  spread(period, Total)  %>%
  inner_join(t.period) 

# tabulate margin totals over status
TOTALS = cbind(Status = 'TOTAL',
                      data.frame( t(sapply(t[,2:ncol(t)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(t)  

t = rbind(t, TOTALS)  # join TOTALS to t

t[is.na(t)] <- 0  # convert NA to 0

kable(t)  # print table

```

\newpage

# Cases by Prefecture

- Y compris les cas confirme et probable, et suspect


```{r gadm_map_data}

# Get GADM rdata files

# see
# http://stackoverflow.com/questions/5126745/gadm-maps-cross-country-comparision-graphics

## load a file from GADM (you just have to specify the countries "special part" of the file name, like "ARG" for Argentina. Optionally you can specify which level you want to have
loadGADM <- function (fileName, level = 1, ...) {
  # load files downloaded from gadm
  load(paste(fileName, "_adm", level, ".RData", sep     = ""))
  gadm
}

## the maps objects get a prefix (like "ARG_" for Argentina)
changeGADMPrefix = function (GADM, prefix) {
  GADM <- spChFIDs(GADM, paste(prefix, row.names(GADM), sep = "_"))
  GADM
}

## load file and change prefix
loadChangePrefix = function (fileName, level = 0, ...) {
  theFile <- loadGADM(fileName, level)
  theFile <- changeGADMPrefix(theFile, fileName)
  theFile
}

## this function creates a SpatialPolygonsDataFrame that contains all maps you specify in "fileNames".
## E.g.: 
## spdf <- getCountries(c("ARG","BOL","CHL"))
## plot(spdf) # should draw a map with Brasil, Argentina and Chile on it.

getCountries = function (fileNames, level = 0, ...) {
  polygon <- sapply(fileNames, loadChangePrefix, level)
  polyMap <- do.call("rbind", polygon)
  polyMap
}


tolerance = .01
country = "GIN"

     gadm = getCountries(country, level=3)
     gadm.f = as.data.frame(gadm) 

    # make list of regions, prefectures, sous prefectures
    regions = gadm.f %>% select(NAME_1,NAME_2,NAME_3) 
    colnames(regions) = c('Region', 'DistrictRes', 'SCRes')

     # fix name of capital in GADM file!
     regions[regions$DistrictRes == 'Conarky', 'DistrictRes'] = 'Conakry'
     regions[regions$Region == 'Conarky', 'Region'] = 'Conakry'

save(regions, file = 'regions.rda')
     
     GIN3 = gadm %>%
#             gSimplify(tolerance, topologyPreserve=T) %>%
            fortify(region = "ID_3") %>%
            mutate( id = as.integer(id)) %>%
            reshape::rename(c("id" = "ID_3")) %>%  #explicit call to rename function in reshape package (as opposed to rename function in plyr package)
            left_join(
              gadm.f[, c("ID_3", "NAME_1", "NAME_2", "NAME_3")])

# fix name of capital in GADM file!
GIN3[GIN3$NAME_2 == 'Conarky', 'NAME_2'] = 'Conakry'

# get location and names for labels
# Admin 3 labels 

          Names = gadm.f %>% 
               select( which(colnames(gadm.f) =="NAME_3"),
                       which(colnames(gadm.f) == "TYPE_3")
               )

          
          GIN3.c =  gadm %>%
                       coordinates()  %>%
                       as.data.frame() %>%
                       mutate(Name = Names[,'NAME_3'],
                              Type = Names[,'TYPE_3']
                              )

# Admin 2

     gadm = getCountries(country, level=2)
     gadm.f = as.data.frame(gadm) 

     GIN2 = gadm %>%
#             gSimplify(tolerance, topologyPreserve=T) %>%
            fortify(region = "ID_2") %>%
            mutate( id = as.integer(id)) %>%
            reshape::rename(c("id" = "ID_2")) %>%
            left_join(
              gadm.f[, c("ID_2", "NAME_1", "NAME_2")])

# fix name of capital in GADM file!
GIN2[GIN2$NAME_2 == 'Conarky', 'NAME_2'] = 'Conakry'

# Admin 2 labels

          Names = gadm.f %>% 
               select( which(colnames(gadm.f) =="NAME_2"),
                       which(colnames(gadm.f) == "TYPE_2")
               )
 
          GIN2.c =  gadm %>%
                       coordinates()  %>%
                       as.data.frame() %>%
                       mutate(Name = Names[,'NAME_2'],
                              Type = Names[,'TYPE_2']
                              ) 

```

```{r theme_defaults}
theme_bare = theme(line = element_blank(),
                   text = element_blank(),
                   line = element_blank(),
                   title = element_blank())

# aesthetic defaults...
color.brewer = "BuPu"  # "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
zero.color = 'white' # add color for cells with zero counts
background.color = 'grey99' 
county.border.color = 'grey70' 
state.border.color = 'grey45'
legend.placement = c(.3, .2)  
legend.font.size = 11 

theme_map =  # coord_map() +
     theme(
          legend.position = legend.placement,
          legend.direction = 'horizontal',
          legend.background = element_blank(),
          legend.title=element_text(size=legend.font.size),
          legend.text = element_text(size=legend.font.size),
          panel.grid.major=element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(colour=background.color, 
                                          fill=background.color),
          panel.background = element_rect(fill=background.color),
          plot.margin = unit(c(0,0,0,0),"cm") # remove plot margins
     )
```

```{r epicurve_facet, fig.height=8, fig.width=6.5, fig.cap='Weekly Case Count by Region, Guinee'}

# epicurve by week, Region (2014 only) 
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()  # collapse to unique values of Region and DistrictRes


# facet by region; fill by prefecture
cases %>% 
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "week")) ) %>%  # set time period
  group_by( Region, DistrictRes, period ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = DistrictRes) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1) ) +
#   scale_fill_manual(values = c('red', 'blue', 'yellow')) +
  facet_grid( Region ~ .)


ggsave('epicurve3.png', width=8, height=6 )

```

```{r tab_district, results='asis'}
p.district = cases %>%  # tabulate margin totals by DistrictRes
  filter(!is.na(period)) %>%
  group_by(DistrictRes) %>% 
  summarize(Total = n()) 

p = cases %>%  # tabulate totals by location
  filter(!is.na(period)) %>%
  group_by( period, DistrictRes ) %>% 
  summarize(Total = n()) %>%
  as.data.frame() %>%  # spread works more reliably on data.frame
  spread(period, Total)  %>%
  merge(p.district) 

p = p %>%
    arrange(DistrictRes )  # sorted by number in must recent week

# tabulate margin totals over status
TOTALS = cbind(DistrictRes = 'TOTAL',
                      data.frame( t(sapply(p[,2:ncol(p)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(p)  

p = rbind(p, TOTALS) # join TOTALS to t

p[is.na(p)] <- 0  # convert NA to 0

kable(p)  # print table

```

```{r prefets, fig.cap="Cas d'Ebola pendant le dernier 21 jours, Guinee"}

# add names of all prefets to case count table
p.name = data.frame(DistrictRes = unique(regions$DistrictRes))
p = merge(p.name, p, all.x=TRUE)
p[is.na(p)] <- 0  # convert NA to 0

# classify values
p$Cases1 =  cut( p[, 'Last 7 days'], 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)

# classify values over past 3 weeks
p$weeks3 = 
    p[, 'Last 7 days'] + p[, 'Previous 8-14 days'] + p[, 'Previous 15-21 days']
      
p$Cases3 =  cut( p[, 'weeks3'], 
                      breaks = c(0,1,10, 20, 100, Inf), 
                      label = c('0','1-9','10-19','20-100', '100 +'),
                      right=FALSE, include.lowest=TRUE)

# Join data with map
p$NAME_2 = p$DistrictRes
map.data.p = merge(GIN2, p, all.x=TRUE, 
                 by.y='NAME_2', by.x = 'NAME_2') %>%
  arrange(order)


# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(5, color.brewer)

# past 3 weeks
g = ggplot( ) + theme_map +
  geom_polygon(data= map.data.p,
               aes(x=long, y=lat, group=group, 
                   fill=Cases3), 
               color="grey50", alpha=1) + 
  geom_text(data= GIN2.c, aes(label = Name, x=V1, y=V2), size=3) +
  scale_fill_manual('Cases', values = colors, drop = FALSE)
g
ggsave('PrefetMap3.png')

```

# Cases by Sous Prefecture

```{r tab_sous_district, results='asis'}

ps.SCRes = cases %>%  # tabulate margin totals by DistrictRes
  filter(!is.na(period)) %>%
  group_by(DistrictRes, SCRes) %>% 
  summarize(Total = n()) 


ps = cases %>%  # tabulate totals by location
  filter(!is.na(period)) %>%
  group_by( period, DistrictRes, SCRes ) %>% 
  summarise(Total = n()) %>%
  as.data.frame() %>%
  spread(period, Total)  %>%
  merge(ps.SCRes) 


# tabulate margin totals over status
TOTALS = cbind(DistrictRes = 'TOTAL', SCRes = '',
                      data.frame( t(sapply(ps[,3:ncol(ps)], sum, na.rm=TRUE) ) )
               )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(ps)  

ps = rbind(ps, TOTALS) # join TOTALS to t

missing.SCRes = ps %>% filter(is.na(SCRes))

ps[is.na(ps)] <- 0  # convert NA to 0

ps[ps$SCRes == '0', 'SCRes'] = 'not_reported'

# kable(ps)  # print table
```

```{r}
# write number of cases with SC not reported.
cat("Sous-prefecture data is missing for", 
    sum(is.na(cases$SCRes)), 
    'records overall, and was not reported for', 
    ps %>% filter(SCRes == 'not_reported') %>% 
                  select(3:5) %>% sum(na.rm = TRUE),
    'of', 
    ps %>% filter(DistrictRes == 'TOTAL') %>% 
                  select(3:5) %>% sum(na.rm = TRUE),
    'during past 3 weeks. \n')
```

```{r sous_prefets_3wks, fig.cap="Cas d'Ebola pendant le dernier 21 jours, Guinee"}

# add names of all prefets to case count table
p.name = data.frame(SCRes = unique(GIN3$NAME_3))
ps = merge(p.name, ps, all.x=TRUE)
ps[is.na(ps)] <- 0  # convert NA to 0

# classify values over past 3 weeks
ps$weeks3 = 
    ps[, 'Last 7 days'] + ps[, 'Previous 8-14 days'] + ps[, 'Previous 15-21 days']
  

# classify values
ps$Cases3 =  cut( ps$weeks3, 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)


# Join data with map
ps$NAME_2 = ps$DistrictRes
ps$NAME_3 = ps$SCRes
map.data.ps = merge(GIN3, ps, all.x=TRUE, 
                 by.y= c('NAME_2', 'NAME_3'), 
                 by.x = c('NAME_2', 'NAME_3')
                 )%>%
  arrange(order)


# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(4, color.brewer)

# last 3 weeks
g = ggplot( ) +  theme_map # + theme_bare 
g = g + theme( axis.text.y = element_blank() )                    
g = g + geom_polygon(data= map.data.ps, 
                     aes(x=long, y=lat, group=group, fill=Cases3),
                     color="grey90", alpha=1)
g = g + geom_path(data= map.data.p, 
                     aes(x=long, y=lat, group=group),
                  color="grey20", alpha=1) 
g = g + geom_text(data= GIN2.c, aes(label = Name, x=V1, y=V2), size=3)
g = g + scale_fill_manual('Cases', values = colors, drop = FALSE)
g

ggsave('sousPrefetMap3.png')
```

```{r sous_prefets_total, fig.cap="Cas d'Ebola depuis le debut d'epidemie, Guinee"}
# add names of all prefets to case count table
p.name = data.frame(SCRes = unique(GIN3$NAME_3))
ps = merge(p.name, ps, all.x=TRUE)
ps[is.na(ps)] <- 0  # convert NA to 0

# classify values
ps$CasesTotal =  cut( ps[, 'Total'], 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)


# Join data with map
ps$NAME_2 = ps$DistrictRes
ps$NAME_3 = ps$SCRes
map.data.ps = merge(GIN3, ps, all.x=TRUE, 
                 by.y= c('NAME_2', 'NAME_3'), 
                 by.x = c('NAME_2', 'NAME_3')
                 )%>%
  arrange(order)



# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(4, color.brewer)

g = ggplot( ) +  theme_map # + theme_bare 
g = g + theme( axis.text.y = element_blank() )                    
g = g + geom_polygon(data= map.data.ps, 
                     aes(x=long, y=lat, group=group, fill=CasesTotal),
                     color="grey90", alpha=1)
g = g + geom_path(data= map.data.p, 
                     aes(x=long, y=lat, group=group),
                  color="grey20", alpha=1) 
g = g + geom_text(data= GIN2.c, aes(label = Name, x=V1, y=V2), size=3)
g = g + scale_fill_manual('Cases', values = colors, drop = FALSE)
g

ggsave('sousPrefetMap_all.png')
```

# Personnel de sante infecté et décedé par suite d'Ebola

```{r hcw_data}

# chart of hcw infection by type of work

h = cases %>% 
  filter( HCW == TRUE) %>% # limit to health care worker (HCW)
  mutate(interval = week(dateOnset), # set time period
         hcw.type = # group types hcw
           ifelse( grepl('LAB', toupper(HCWposition)), 'Lab',
              ifelse( grepl('M.D', toupper(HCWposition)),
                           'Medecin', 
                ifelse( grepl('INF', toupper(HCWposition))
                                   , 'Infirmiere', 
                                  'Other')))
         ) %>%  
  filter( year(dateOnset) == 2014 ) %>%
  merge(regions, all.x = TRUE) 
```

Since the onset of the epidemic, there have been `r nrow(h) ` cases among health care workers.

```{r chart_hcw_data, fig.height=8, fig.width=6.5}
h %>%
  ggplot( aes(x = interval ) )+
  geom_bar( aes(y = ..count.., fill = hcw.type), 
            binwidth=1, color = 'white') +
#   geom_text( aes(label = StatusAsOfCurrentDate) ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = 'top' ) +
  scale_x_discrete('Week') +
  scale_fill_discrete('Position') +
  scale_shape_discrete('Status') +
  facet_grid( DistrictRes ~ .) 

ggsave('epicurve_HCW.png', width=8, height=6 )

```

For the sake of visualiztion, data avaialable on the type health work was classified as Infirmiere, Lab, Medecin, or other.  

The following table lists the position descriptions available for type of health care worker.  

`r kable(h %>% count(hcw.type))`

The free text given for each type is shown below.  


Infirmiere: 
```{r}
I = h  %>%  
  filter( year(dateOnset) == 2014 & hcw.type %in% 'Infirmiere')

paste( unique(I$HCWposition), collapse = ', ') 
```


Lab: 
```{r}
L = h  %>%  
  filter( year(dateOnset) == 2014 & hcw.type %in% 'Lab')

paste( unique(L$HCWposition), collapse = ', ') 
```


Medecin: 
```{r}
M= h  %>%  
  filter( year(dateOnset) == 2014 & hcw.type %in% 'Medecin')

paste( unique(M$HCWposition), collapse = ', ') 
```

Other: 
```{r}
O = h  %>%  
  filter( year(dateOnset) == 2014 & hcw.type %in% 'Other')

paste( unique(O$HCWposition), collapse = ', ') 
```
