---
title: "VHF DATA"
author: "EPI TASK FORCE"
date: "Tuesday, November 18, 2014"
output:
  slidy_presentation: default
  ioslides_presentation:
    transition: faster
    widescreen: yes
  beamer_presentation: default
---

```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }
options(tz="GMT+0")
Sys.setenv(TZ='GMT+0')

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warnings=FALSE, results='asis',
                cache = FALSE)
library(mi)
library(Hmisc)
library(XLConnect, quietly=TRUE)
library(lubridate)
library(plyr)

library(ggplot2)
library(scales)
library(RColorBrewer)

library(tidyr)
library(dplyr, quietly=TRUE) # order is important...

library(rgeos)
```

# VHF Data Elements

## VHF Data, Guinea: **Cases**

```{r read_cas_G}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.
### note need to correct cell  AW-1455 from '=+Yambala' to 'Yambala'

dir = 'fieldData/'

file.name = "guinea_cases_10_Nov"  

excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
case_data.guinea = readWorksheet(wb, file.name, )

# csv.file = paste0(dir, file.name, '.csv')
# case_data = read.csv(csv.file, header=TRUE)

save(case_data.guinea, file='case_data.guinea.rda')
```

```{r Define_case_data_G}
load('case_data.guinea.rda')

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Cas','Confirme',
                                          'Probable','Suspect', 'Exclude'))

# if else with dates can be troublesome.  see sol at 
# http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no){ class.y <- class(yes)
                                  X <- ifelse(cond,yes,no)
                                  class(X) <-class.y; return(X)}


last.report.date = max(case_data.guinea$DateReport, na.rm=T) # last day with report
current.week = week(last.report.date)  

cases.guinea = case_data.guinea %>% 
  select( ID, StatusAsOfCurrentDate, 
          EpiCaseDef , StatusReport, FinalStatus,
          ThisCaseIsAlsoContact, TotalContactsListed, 
          DateReport, DateOnset,
          DateHospitalCurrentAdmit, DateIsolationCurrent,
          Age, AgeUnit, Gender, DateDeath, 
          VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, HCWposition, HCWFacility, 
          Funeral, Animals, 
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EpiCaseDef==4) %>% # drop 'excluded'
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  inner_join(epiCaseDef) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          CaseMonth = paste(month(dateOnset), year(dateOnset)),
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirme','Probable', 'Suspect'))

save(cases.guinea, file = 'cases.guinea.rda')
```

```{r Guinea_missing, fig.height=5}
load('cases.guinea.rda')

data = cases.guinea %>% 
#   arrange(DateReport) %>%
  select(EpiCaseDef,  DateReport, 
         DateOnset, DateIsolationCurrent, 
         DateHospitalCurrentAdmit,
         StatusAsOfCurrentDate, 
         Age , Gender, 
         DistrictRes, SCRes,
         HCW, Funeral, Animals, 
         Contact) 

par(mai=c(1,3,0,0.1))
missing.pattern.plot( data ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(data), ")" ),
                      main = ''
                      )

# cat("VHF cases reports through", last.report.date )

# mis =  mi.info(cases2) 
# miss = data.frame(mis$name, mis$nmis, mis$type)
# colnames(miss) = c('Variable', 'Nmiss', 'Type')
# kable(miss, row.names=F)

```

Available data (blue) for select variable among confirmed, suspect, and probable cases

## VHF Data, Sierra Leone: **Cases**

```{r read_cas_SL}

# In order for this import to work, must do 2 things first:
# 1. Open accdb each VHF_CIF file and export to EXCEL, then
# 2. Export to CSV

dir = 'fieldData/'

file.name = "VHF_CIF1"
cif1 = read.csv( paste0(dir, file.name, '.csv'), header = TRUE, stringsAsFactors = FALSE)

file.name = "VHF_CIF2"
cif2 = read.csv( paste0(dir, file.name, '.csv'), header = TRUE, stringsAsFactors = FALSE)

case_data.sl = inner_join(cif1, cif2) %>%
  mutate( DateReport = dmy(DateReport),
          DateOnset = dmy(DateOnset) ,
          DateIsolationCurrent = dmy(DateIsolationCurrent))

save(case_data.sl, file='case_data.sl.rda')
```

```{r Define_case_data_SL}
load('case_data.sl.rda')

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Case','Confirmed',
                                          'Probable','Suspect', 'Excluded'))

# if else with dates can be troublesome.  see sol at 
# http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no){ class.y <- class(yes)
                                  X <- ifelse(cond,yes,no)
                                  class(X) <-class.y; return(X)}

last.report.date = max(case_data.sl$DateReport, na.rm=T) # last day with report
current.week = week(last.report.date)  

cases.sl = case_data.sl %>% 
  select( ID, StatusAsOfCurrentDate, 
          EPICaseDef , StatusReport, FinalStatus,
          DateReport, DateOnset,
          DateHospitalCurrentAdmit, DateIsolationCurrent,
          Age, AgeUnit, Gender, DateDeath, 
          VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, HCWposition, HCWFacility, 
          Funeral, Animals, 
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EPICaseDef==4) %>% # drop 'excluded'
  mutate( EPICaseDef = as.integer(EPICaseDef)) %>%
  inner_join(epiCaseDef, by = c('EPICaseDef' = 'EpiCaseDef')) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          CaseMonth = paste(month(dateOnset), year(dateOnset)),
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirmed','Probable', 'Suspect') & 
              dateOnset > ymd('2014-JUN-01'))

# function to convert text to Title Case (when strict = F, ignores all upper case words)
TitleCase = function(s, strict = TRUE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                  {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

# TitleCase(c("using AIC for model selection"))
# TitleCase(c("HELP"))

cases.sl = cases.sl %>%
  mutate(
    DistrictRes = TitleCase(DistrictRes),
    SCRes = TitleCase(SCRes)
    )

# table(cases$DistrictRes)

  cases.sl[cases.sl$DistrictRes %in% 'Western Area Rural', 'DistrictRes'] = 'Western Rural'
  
  cases.sl[cases.sl$DistrictRes %in% "Western Area Urban", 'DistrictRes'] = 'Western Urban'
  
  cases.sl[cases.sl$DistrictRes %in% 'Tonkllili', 'DistrictRes'] = 'Tonkolili'

    cases.sl[cases.sl$DistrictRes %in% 'Free Town', 'DistrictRes'] = 'Western Urban'

    cases.sl[cases.sl$DistrictRes %in% '..tonkolili', 'DistrictRes'] = 'Tonkolili'

  cases.sl[cases.sl$DistrictRes %in% 'Kenenma', 'DistrictRes'] = 'Kenema'

# save cases data
save(cases.sl, file='cases.sl.rda')

```

```{r SL_missing, fig.height=5}
load('cases.sl.rda')

data = cases.sl %>% select(EPICaseDef,  
                        DateOnset, DateIsolationCurrent, 
                        StatusAsOfCurrentDate, 
                        Age , Gender, 
                        DistrictRes, SCRes,
                        HCW, Funeral, Animals, 
                        Contact)

par(mai=c(1,3,0,0))
missing.pattern.plot( data ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(data), ")" ),
                      main = ''
                      )

```

Available data (blue) for select variable among confirmed, suspect, and probable cases

## VHF Data, Guinea: **Contacts**

```{r read_contacts}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.

file.name = "guinea_contacts_10_Nov"

dir = 'fieldData/'

excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
contact_data.guinea = readWorksheet(wb, file.name)

contacts.guinea = contact_data.guinea %>%
  mutate(ContactMonth = paste(month(DateLastContact), year(DateLastContact)) ) %>%
  arrange(DateLastContact)

save(contacts.guinea, file = "contacts.guinea.rda")

```

```{r mi_contacts, fig.height=5}
# paste(colnames(contacts.guinea), collapse = ', ')

contact.cols = c('ThisContactIsAlsoCase', 'Surname', 'OtherNames', 'Gender', 'Age', 'HeadHousehold', 'Village', 'District', 'SubCounty', 'SourceCaseID', 'SourceCase', 'RelationshipToCase', 'ContactTypes', 'DateLastContact', 'DateOfLastFollowUp', 'FinalOutcome')

par(mai=c(1,3,0,.1))
missing.pattern.plot( contacts.guinea[, contact.cols] ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(contacts.guinea), ")" ),
                      main = '')
```

- Approximately 3 contacts per case, evidence of missing contact reports

## VHF Data, Sierra Leone: **Contacts**

- no data available


# Suggested Process Measures

## Number of Contacts Traced
```{r contacts_traced_g, fig.height=3}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

load('contacts.guinea.rda')

data = contacts.guinea %>% 
  select( DateLastContact, RelationshipToCase, District, DateOfLastFollowUp, FinalOutcome) %>%
  merge(Region, all.x = TRUE, by.x ='District' , by.y = 'DistrictRes' ) %>%
  mutate( daysFollowed = as.numeric( as.Date(DateOfLastFollowUp) - as.Date(DateLastContact) ),
          RelationshipToCase = factor(RelationshipToCase),
          month = month(DateLastContact)
          ) 

data. = data %>% count(Region, month) 

data. = data. %>% filter(Region %in% c('Conakry'))
 
g = ggplot(data., aes( month, n)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Number Contacts') +
  geom_bar(stat = 'identity') +
  facet_grid( Region ~ .)
# g

load('cases.guinea.rda')
case.month = cases.guinea %>% mutate( month = month(dateOnset) ) %>% 
  merge(Region, all.x = TRUE, by = 'DistrictRes') %>%
  filter(Region %in% 'Conakry' | is.na(Region)) %>%
  count(month)

contact.case = data. %>% inner_join(case.month, by='month') %>%
  mutate(contact.case = round(n.x/n.y, 1))

h = ggplot(data=contact.case, aes(month, contact.case)) + geom_line() +
  scale_y_continuous('Contacts per Case' ) +
  scale_x_continuous('Month', breaks = c(1:12))

library(gridExtra)
grid.arrange(g, h)
```

- VHF only has contacts from Conakry
- Other contact lists maintained on offline spreadsheets
- Mid-October, Conakry contacts moved to offline spreadsheet only
- Actual number of contacts:case ~ 20:1

## Cases that were Contacts, Sep-Oct, Conakry

```{r cases_found_among_contact, fig.height=4, fig.width=10}

load('cky.cases.rda')

sourceCases = cases %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( 
          period = as.Date(floor_date(dateOnset, "week")),
          SCRes = toupper(SCRes)
          ) %>% 
  group_by( period ) %>%
  summarize( n = n(),
             contact = sum(WasContact, na.rm=T),
             caseContacts = contact/n
             )  %>%
  ## plot
  ggplot( aes(x = period , y = caseContacts) ) +
  geom_line() +  geom_point( aes(size = n)) + 
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Percent Cases Known Contacts') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) 

# percent cases that were contacts
x = cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  summarise( n = n(),
             contact = sum(WasContact),
             caseContacts = percent(contact/n)
             )

```

- Probablisitic linkage of contact registry with case registry
- Overall, `r x[1,3]` of cases were known contacts
- Limited by missing contact records

## Cases that were Contacts, Sep-Oct, Conakry

```{r epicurve_cases_were_contacts, fig.height=6, fig.width=10}

load('cky.cases.rda')

# t = with(cky.cases, table(SCRes, WasContact))
# kable( t  )

# cat('NOTE: For', sum(is.na(cky.cases$SCRes)) ,'cases, DistrictRes is missing')

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

sourceCases = cases %>% 
  filter(ID %in% cky.cases$SourceCaseID) %>%
  rename(SourceCaseID = ID, sourceSCRes = SCRes) %>%
  select(SourceCaseID, sourceSCRes) 
  
cky.cases %>% 
  left_join(sourceCases, by= c('SourceCaseID'='SourceCaseID')) %>%
  mutate( idnum = substrRight(ID, 4),
          period = as.Date(floor_date(dateOnset, "week")),
          SCRes = toupper(SCRes),
          VillageRes = toupper(VillageRes),
          WasContact = factor(WasContact, c(0,1), c('No','Yes'))
          ) %>% 
  group_by( period, SCRes, idnum, WasContact, sourceSCRes ) %>%
  summarize( freq = n() ) %>%  
  mutate( cumsum = cumsum(freq),
          midpoint = (cumsum(freq) - 0.5 * freq)) %>%
  ## plot
  ggplot( aes(x = period , y = freq) ) +
  geom_bar( stat = 'identity', aes( fill=sourceSCRes), color = "grey90" ) +
  scale_fill_discrete('Case Was a Contact of Case in:') + # values = c('grey90', 'black')) +
  scale_color_discrete('Case Was a Contact of Case in:') +
#   geom_text( aes(y = midpoint, ymax=midpoint, label = idnum), size = 2, hjust=1) +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 2),
       legend.position = 'top' ) +
  facet_grid(SCRes ~ .)
```

## Secondary Cases, Sep-Oct, Conakry

```{r secondary_cases, fig.height=4, fig.width=10}
load('cky.cases.rda')

john.ben.interval = new_interval(ymd("2014-SEP-1"), ymd("2014-OCT-30"))

t = cky.cases %>% 
  filter(dateOnset %within% john.ben.interval ) %>%
  count(SourceCaseID) %>%
  arrange(-n) %>%
  rename(Number_2ndary_Cases = n)

# kable(t)

total.secondary = sum(t$Number_2ndary_Cases, na.rm=T)
number.sources = sum(!is.na(t$SourceCaseID))

t %>% filter(!is.na(t$SourceCaseID)) %>%
ggplot(aes(x=Number_2ndary_Cases )) +
  geom_bar() +
  scale_x_continuous('Number of secondary cases', 
                     breaks = seq(1, max(t$Number_2ndary_Cases), 2)
                     ) 
#   geom_annotate(x=7, y=10, type='text', text=paste('))
```

- There were `r x[1,2]` (`r percent(x[1,2] / nrow(cky.contacts) )`) cases among `r nrow(cky.contacts) ` contacts.
- One case resulted in 15 cases among `r cky.contacts %>% filter(SourceCaseID %in% 'GUI-CKY-14-1365') %>% count()` contacts.  
- One case resulted in 12 cases among `r cky.contacts %>% filter(SourceCaseID %in% 'GUI-CKY-14-0988') %>% count()` contacts.


# Other Indicators


# Days to Isolation

## Days to Isolation, Guinea
```{r isolation_g, warning=FALSE}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28)
 
g = ggplot(data, aes( month(DateOnset), daysToIso)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("red", "green")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  
# + 
#   facet_grid( Region ~ .)
g
```

## Days to Isolation by Region, Guinea
### Conakry v. Forrest Region
```{r isolation_g_region, warning=FALSE}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28) 

data = data %>% filter(Region %in% c('Nzérékoré', 'Conakry'))
 
g = ggplot(data, aes( month(DateOnset), daysToIso)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("red", "green")) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  +
  facet_grid( Region ~ .)
g
```

## Days to Isolation, SL
### Bombali, Kailahun, Kenema, Moyamba, Port Loko, Tonkolili 
```{r isolation_sl_region, warning=FALSE}

# Time from onset to isolation
load('cases.sl.rda')
data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 20) 

data = data %>% filter(DistrictRes %in% c('Bombali', 'Kailahun', 'Kenema', 'Moyamba',' Port Loko', 'Tonkolili' ))

g = ggplot(data, aes( month(DateOnset), daysToIso)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Isolation', breaks = seq(0, 21, 2)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c("grey", "green", 'red')) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA)  +
  facet_grid( DistrictRes ~ .)
g
```


# Survival

## Survival by Days to Isolation, Guinea
### Conakry v. Forrest Region
```{r survival_g, warning=FALSE, fig.height=4}
load('regions.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()

# Time from onset to isolation
load('cases.guinea.rda')
data = cases.guinea %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28) 

data = data %>% group_by(daysToIso, Region) %>% 
  summarise(n=n(),
            dead = sum(ifelse(StatusAsOfCurrentDate %in% 'Décédé',1,0)),
            alive = sum(ifelse(StatusAsOfCurrentDate %in% 'Vivant',1,0))
            ) %>%
  mutate(survival = alive/ (dead + alive))

data = data %>% filter(Region %in% c('Nzérékoré', 'Conakry'))
 
g = ggplot(data, aes(x = daysToIso, y = survival, label = n)) +
    geom_smooth(alpha = 0) +
    geom_point(aes(size = n)) +
#     geom_text(vjust = 11) +
    scale_x_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    scale_y_continuous( labels = percent, breaks = seq(0,1,.1), limits = c(0,1.1) ) +
    scale_size_continuous('Patients') +
  facet_grid( Region ~ .)
g
```

- Conakry VHF data likely over-represents survival because status updates delayed/missing

## Survival, Donka ETU (Conakry)

```{r donka, warning=FALSE}
# library(XLConnect)
wb = loadWorkbook('fieldData/GUI_Conakry_EbLinelist_2014_11_14.xlsx')
donka = readWorksheet(wb, 'registre') %>% 
  filter( Statut == 'Confirmé') 

donka.= donka %>% 
          select( Num, Nom..Prénom..âge..genre, Age, Genre, Commentaires, Date.Symptômes, Sorties, Durée.de.séjour, Origine) %>%
  arrange( Date.Symptômes ) %>%
  mutate( month = month(Date.Symptômes)) %>%
  group_by( month ) %>%
  summarize( n = n(),
             died = sum(ifelse(Sorties %in% 'Décédé',1,0)) ,
             survival = died / n
             )

g = ggplot(donka., aes(month, survival)) + 
  geom_point(aes(size = n)) + geom_line() +
    scale_x_continuous('Month', breaks = 1:12) +
    scale_y_continuous( labels = percent, breaks = seq(0,1,.1), limits = c(0,1) ) +
    scale_size_continuous('Patients') 
g
```

- Restricted to confirmed cases admitted to Donka ETU
- Mean survival rate = `r percent( nrow(donka %>% filter(Sorties %in% 'Décédé')) / nrow(donka)) `

## Survival, Donka ETU (Conakry), by Case-Origin

```{r donka_origin,  warning=FALSE}
# library(XLConnect)
wb = loadWorkbook('fieldData/GUI_Conakry_EbLinelist_2014_11_14.xlsx')
donka = readWorksheet(wb, 'registre') %>% 
  filter( Statut == 'Confirmé') 

# clean district names with fuzz matching (agrepl)
donka$origin = NA
Origine = 'RATOMA'
donka[ agrepl(Origine, donka$Origine, ignore.case = T), 'origin' ] = Origine
Origine = 'MATOTO'
donka[ agrepl(Origine, donka$Origine, ignore.case = T), 'origin' ] = Origine
Origine = 'MATAM'
donka[ agrepl(Origine, donka$Origine, ignore.case = T), 'origin' ] = Origine
Origine = 'KALOUM'
donka[ agrepl(Origine, donka$Origine, ignore.case = T), 'origin' ] = Origine
Origine = 'DIXINN'
donka[ agrepl(Origine, donka$Origine, ignore.case = T), 'origin' ] = Origine

donka[is.na(donka$origin), 'origin' ] = 'Not Conakry'
# table(donka$origin, useNA = 'always')

donka.= donka %>% 
          select( Num, Nom..Prénom..âge..genre, Age, Genre, Commentaires, Date.Symptômes, Sorties, Durée.de.séjour, Origine, origin) %>%
  arrange( Date.Symptômes ) %>%
  mutate( month = month(Date.Symptômes)) %>%
  group_by( month, origin ) %>%
  summarize( n = n(),
             died = sum(ifelse(Sorties %in% 'Décédé',1,0)) ,
             survival = died / n
             )

g = ggplot(donka., aes(month, survival)) + 
  geom_point(aes(size = n)) + geom_line() +
    scale_x_continuous('Month', breaks = 1:12) +
    scale_y_continuous( labels = percent, breaks = seq(0,1,.2), limits = c(0,1) ) +
    scale_size_continuous('Patients')  +
  facet_grid( origin ~ .)
g

```

- Restricted to confirmed cases admitted to Donka ETU

## Survival by Days to Isolation, Sierra Leone
```{r survival_sl, warning=FALSE}

# Time from onset to isolation
load('cases.sl.rda')

data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28) 

data = data %>% group_by(daysToIso) %>% 
  summarise(n=n(),
            dead = sum(ifelse(StatusAsOfCurrentDate %in% 'DEAD',1,0)),
            alive = sum(ifelse(StatusAsOfCurrentDate %in% 'ALIVE',1,0))
            ) %>%
  mutate(survival = alive/ (dead + alive))
 
g = ggplot(data, aes(x = daysToIso, y = survival, label = n)) +
    geom_smooth(alpha = 0) +
    geom_point(aes(size = n)) +
#     geom_text(vjust = 11) +
    scale_x_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    scale_y_continuous( labels = percent, breaks = seq(0,1,.1), limits = c(0,1.1) ) +
    scale_size_continuous('Patients') 
g
```
 
## Survival by Days to Isolation, SL
### Bombali, Kailahun, Kenema, Moyamba, Port Loko, Tonkolili 
```{r survival_sl_region, warning=FALSE}

load('cases.sl.rda')

data = cases.sl %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28) 

data = data %>% filter(DistrictRes %in% c('Bombali', 'Kailahun', 'Kenema', 'Moyamba',' Port Loko', 'Tonkolili' ))

data = data %>% group_by(daysToIso, DistrictRes) %>% 
  summarise(n=n(),
            dead = sum(ifelse(StatusAsOfCurrentDate %in% 'DEAD',1,0)),
            alive = sum(ifelse(StatusAsOfCurrentDate %in% 'ALIVE',1,0))
            ) %>%
  mutate(survival = alive/ (dead + alive))
 
g = ggplot(data, aes(x = daysToIso, y = survival, label = n)) +
    geom_smooth(alpha = 0) +
    geom_point(aes(size = n)) +
#     geom_text(vjust = 11) +
    scale_x_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    scale_y_continuous( labels = percent, breaks = seq(0,1,.1), limits = c(0,1.1) ) +
    scale_size_continuous('Patients') +
  facet_grid(DistrictRes ~ .)
g
```
 
## StatusAsOfCurrentDate variable, Bombali
Something seems wrong--nearly 100% survival!

```{r Bombali_error}
load('cases.sl.rda')

d = cases.sl %>% filter(DistrictRes %in% 'Bombali') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
  filter(daysToIso < 28) %>%
  count(StatusAsOfCurrentDate)

kable(d)
```

## StatusAsOfCurrentDate variable, Bombali

### same data set with no date restriction

```{r Bombali_error2}
load('cases.sl.rda')

d = cases.sl %>% filter(DistrictRes %in% 'Bombali') %>%
#   filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) %>%
#   filter(daysToIso < 28) %>%
  count(StatusAsOfCurrentDate)

kable(d)
```

- many records had missing Days to Isolation because DateIsolationCurrent < DateOnset 
- when this restriction removed, many more patients, most of whom died
- Likely that some data entry used improper date format (American vs European)
