---
title: "Sierra Leone"
author: "CDC/OMS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6.5
    
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes

header-includes:
- \usepackage[section]{placeins}
- \usepackage{setspace}
- \usepackage{relsize}
---

```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }
options(tz="GMT+0")
Sys.setenv(TZ='GMT+0')

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warning=FALSE, results='asis')

library(XLConnect, quietly=TRUE)
library(lubridate)
library(reshape)  # use for rename() function

library(rgeos)
library(sp)

suppressPackageStartupMessages(library(ggplot2))
library(scales)
library(grid)
library(RColorBrewer)

# library(devtools)
# devtools::install_github("hadley/tidyr")
library(tidyr) # install from github (CRAN version not working...)
library(dplyr, quietly=TRUE) # order is important...

```

# Data 

- Data loaded from export of VHF database.^[export cases as .csv, then open and save file as excel]

```{r read_cas}

# In order for this import to work, must do 2 things first:
# 1. Open accdb each VHF_CIF file and export to EXCEL, then
# 2. Export to CSV

dir = 'fieldData/'

file.name = "VHF_CIF1"
cif1 = read.csv( paste0(dir, file.name, '.csv'), header = TRUE, stringsAsFactors = FALSE)

file.name = "VHF_CIF2"
cif2 = read.csv( paste0(dir, file.name, '.csv'), header = TRUE, stringsAsFactors = FALSE)

case_data = inner_join(cif1, cif2) %>%
  mutate( DateReport = dmy(DateReport),
          DateOnset = dmy(DateOnset) ,
          DateIsolationCurrent = dmy(DateIsolationCurrent))

```

```{r Define_case_data}

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Case','Confirmed',
                                          'Probable','Suspect', 'Excluded'))

# if else with dates can be troublesome.  see sol at 
# http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no){ class.y <- class(yes)
                                  X <- ifelse(cond,yes,no)
                                  class(X) <-class.y; return(X)}

last.report.date = max(case_data$DateReport, na.rm=T) # last day with report
current.week = week(last.report.date)  

cases = case_data %>% 
  select( ID, StatusAsOfCurrentDate, 
          EPICaseDef , StatusReport, FinalStatus,
          DateReport, DateOnset,
          DateHospitalCurrentAdmit, DateIsolationCurrent,
          Age, AgeUnit, Gender, DateDeath, 
          VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, HCWposition, HCWFacility, 
          Funeral, Animals, 
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EPICaseDef==4) %>% # drop 'excluded'
  mutate( EPICaseDef = as.integer(EPICaseDef)) %>%
  inner_join(epiCaseDef, by = c('EPICaseDef' = 'EpiCaseDef')) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          CaseMonth = paste(month(dateOnset), year(dateOnset)),
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirmed','Probable', 'Suspect') & 
              dateOnset > ymd('2014-JUN-01'))

```

```{r Clean_Locations}

# function to convert text to Title Case (when strict = F, ignores all upper case words)
TitleCase = function(s, strict = TRUE) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                  {s <- substring(s, 2); if(strict) tolower(s) else s},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

# TitleCase(c("using AIC for model selection"))
# TitleCase(c("HELP"))

cases = cases %>%
  mutate(
    DistrictRes = TitleCase(DistrictRes),
    SCRes = TitleCase(SCRes)
    )

# table(cases$DistrictRes)

  cases[cases$DistrictRes %in% 'Western Area Rural', 'DistrictRes'] = 'Western Rural'
  
  cases[cases$DistrictRes %in% "Western Area Urban", 'DistrictRes'] = 'Western Urban'
  
  cases[cases$DistrictRes %in% 'Tonkllili', 'DistrictRes'] = 'Tonkolili'

    cases[cases$DistrictRes %in% 'Free Town', 'DistrictRes'] = 'Western Urban'

    cases[cases$DistrictRes %in% '..tonkolili', 'DistrictRes'] = 'Tonkolili'

  cases[cases$DistrictRes %in% 'Kenenma', 'DistrictRes'] = 'Kenema'

# save cases data
save(cases, file='casesSL.rda')

```

```{r missing_date, results='asis'}
#  list records missing period
missingPeriod = cases %>% filter(is.na(dateOnset)) %>% select(ID)
```

- The most recent case included in this report had symptom onset on 
**`r format(max(cases$DateReport, na.rm=TRUE), "%d-%B-%Y")`**.

- There are `r nrow(missingPeriod)` records with missing dates:   
`r cat(missingPeriod$ID, sep=', ')`

# Epicurve by Month

```{r epicurve_monthly, fig.cap='Monthly Ebola Cases, Guinee'}

# epicurve by month (2014 only) 

cases %>% 
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "month")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('month', labels = date_format("%b-%dd"),
               breaks = date_breaks("month")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve0_SL.png')
  

```

# Epicurves By Week (by report date and onset date)

```{r epicurves_weekly_DateReported, fig.height=3.5, fig.cap='Weekly Ebola Cases, Guinee (by date reported)'}

cases %>% 
  filter( year(DateReport) == 2014 ) %>%
  arrange(DateReport) %>%
  mutate( interval = week(DateReport),
          period = as.Date(floor_date(DateReport, "week")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve1_SL.png')
```

```{r epicurves_weekly_onsetDate, fig.height=3.5, fig.cap='Weekly Ebola Cases, Guinee (by onset date)'}
cases %>% 
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "week")) ) %>%  # set time period
  group_by( period, Status ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = Status) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week beginning', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
       legend.position = 'top' ) +
  scale_fill_manual(values = c('red', 'blue', 'yellow')) 

ggsave('epicurve2_SL.png')

```

# Case-Status by Week

```{r tab_status_byOnsetWeek}

t.period_wk = cases %>%    # tabulate margin totals by status
  filter(!is.na(period_wk)) %>%
  group_by(Status) %>% 
  summarise(Total = n())

t = cases %>%  # tabulate totals by status and period_wk
  filter(!is.na(period_wk)) %>%
  group_by(Status, period_wk) %>% 
  summarise(Total = n()) %>% 
  as.data.frame() %>%  # spread works more reliably on data.frame
#   mutate( Percent = percent(Total/sum(Total, na.rm=TRUE)) ) %>%
  spread(period_wk, Total)  %>%
  inner_join(t.period_wk) 

# tabulate margin totals over status
TOTALS = cbind(Status = 'TOTAL',
                      data.frame( t(sapply(t[,2:ncol(t)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(t)  

t = rbind(t, TOTALS)  # join TOTALS to t

t[is.na(t)] <- 0  # convert NA to 0

# Calculate number of days reported in current week
daysThisWeek = cases %>% 
  select(period_wk, dateOnset) %>% 
  filter(
    if(sum(period_wk %in% 'Week')>0){period_wk=='Week'}else{period=='Week_1'}
       ) %>% 
  summarise( 
    days = max(dateOnset)- min(dateOnset) + 1 
    )

cat('The current week has', daysThisWeek$days, 'days\n')

kable(t)  # print table

```

# Case-Status by 7-day period, begining with last reported case

```{r tab_status_7days}

t.period = cases %>%    # tabulate margin totals by status
  filter(!is.na(period)) %>%
  group_by(Status) %>% 
  summarise(Total = n())

t = cases %>%  # tabulate totals by status and period
  filter(!is.na(period)) %>%
  group_by(Status, period) %>% 
  summarise(Total = n()) %>% 
  as.data.frame() %>%  # spread works more reliably on data.frame
#   mutate( Percent = percent(Total/sum(Total, na.rm=TRUE)) ) %>%
  spread(period, Total)  %>%
  inner_join(t.period) 

# tabulate margin totals over status
TOTALS = cbind(Status = 'TOTAL',
                      data.frame( t(sapply(t[,2:ncol(t)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(t)  

t = rbind(t, TOTALS)  # join TOTALS to t

t[is.na(t)] <- 0  # convert NA to 0

kable(t)  # print table

```

\newpage
cases by Prefecture

- Y compris les cas confirme et probable, et suspect


```{r gadm_map_data, cache=TRUE}

# Get GADM rdata files

# see
# http://stackoverflow.com/questions/5126745/gadm-maps-cross-country-comparision-graphics

## load a file from GADM (you just have to specify the countries "special part" of the file name, like "ARG" for Argentina. Optionally you can specify which level you want to have
loadGADM <- function (fileName, level = 1, ...) {
  # load files downloaded from gadm
  load(paste(fileName, "_adm", level, ".RData", sep     = ""))
  gadm
}

## the maps objects get a prefix (like "ARG_" for Argentina)
changeGADMPrefix = function (GADM, prefix) {
  GADM <- spChFIDs(GADM, paste(prefix, row.names(GADM), sep = "_"))
  GADM
}

## load file and change prefix
loadChangePrefix = function (fileName, level = 0, ...) {
  theFile <- loadGADM(fileName, level)
  theFile <- changeGADMPrefix(theFile, fileName)
  theFile
}

## this function creates a SpatialPolygonsDataFrame that contains all maps you specify in "fileNames".
## E.g.: 
## spdf <- getCountries(c("ARG","BOL","CHL"))
## plot(spdf) # should draw a map with Brasil, Argentina and Chile on it.

getCountries = function (fileNames, level = 0, ...) {
  polygon <- sapply(fileNames, loadChangePrefix, level)
  polyMap <- do.call("rbind", polygon)
  polyMap
}


tolerance = .01
country = "SLE"

     gadm = getCountries(country, level=3)
     gadm.f = as.data.frame(gadm) 

    # make list of regions, prefectures, sous prefectures
    regions = gadm.f %>% select(NAME_1,NAME_2,NAME_3) 
    colnames(regions) = c('Region', 'DistrictRes', 'SCRes')

     # fix name of capital in GADM file!
#      regions[regions$DistrictRes == 'Conarky', 'DistrictRes'] = 'Conakry'
#      regions[regions$Region == 'Conarky', 'Region'] = 'Conakry'

save(regions, file = 'regionsSL.rda')

```

```{r gadm_map_data2}

load('regionsSL.rda')

     SLE3 = gadm %>%
#             gSimplify(tolerance, topologyPreserve=T) %>%
            fortify(region = "ID_3") %>%
            mutate( id = as.integer(id)) %>%
            reshape::rename(c("id" = "ID_3")) %>%  #explicit call to rename function in reshape package (as opposed to rename function in plyr package)
            left_join(
              gadm.f[, c("ID_3", "NAME_1", "NAME_2", "NAME_3")])

# fix name of capital in GADM file!
# SLE3[SLE3$NAME_2 == 'Conarky', 'NAME_2'] = 'Conakry'

# get location and names for labels
# Admin 3 labels 

          Names = gadm.f %>% 
               select( which(colnames(gadm.f) =="NAME_3"),
                       which(colnames(gadm.f) == "TYPE_3")
               )

          
          SLE3.c =  gadm %>%
                       coordinates()  %>%
                       as.data.frame() %>%
                       mutate(Name = Names[,'NAME_3'],
                              Type = Names[,'TYPE_3']
                              )

# fix name of capital in GADM file!
# SLE3.c[SLE3.c$Name == 'Conarky', 'Name'] = 'Conakry'

# Admin 2

     gadm = getCountries(country, level=2)
     gadm.f = as.data.frame(gadm) 

     SLE2 = gadm %>%
#             gSimplify(tolerance, topologyPreserve=T) %>%
            fortify(region = "ID_2") %>%
            mutate( id = as.integer(id)) %>%
            reshape::rename(c("id" = "ID_2")) %>%
            left_join(
              gadm.f[, c("ID_2", "NAME_1", "NAME_2")])

# fix name of capital in GADM file!
# SLE2[SLE2$NAME_2 == 'Conarky', 'NAME_2'] = 'Conakry'

# Admin 2 labels

          Names = gadm.f %>% 
               select( which(colnames(gadm.f) =="NAME_2"),
                       which(colnames(gadm.f) == "TYPE_2")
               )
 
          SLE2.c =  gadm %>%
                       coordinates()  %>%
                       as.data.frame() %>%
                       mutate(Name = Names[,'NAME_2'],
                              Type = Names[,'TYPE_2']
                              ) 

# fix name of capital in GADM file!
# SLE2.c[SLE2.c$Name == 'Conarky', 'Name'] = 'Conakry'

```

```{r theme_defaults}
theme_bare = theme(line = element_blank(),
                   text = element_blank(),
                   line = element_blank(),
                   title = element_blank())

# aesthetic defaults...
color.brewer = "BuPu"  # "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
zero.color = 'white' # add color for cells with zero counts
background.color = 'grey99' 
county.border.color = 'grey70' 
state.border.color = 'grey45'
legend.placement = c(.5, -.05)  
legend.font.size = 11 

theme_map =  # coord_map() +
     theme(
          legend.position = legend.placement,
          legend.direction = 'horizontal',
          legend.background = element_blank(),
          legend.title=element_text(size=legend.font.size),
          legend.text = element_text(size=legend.font.size),
          panel.grid.major=element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(colour=background.color, 
                                          fill=background.color),
          panel.background = element_rect(fill=background.color),
          plot.margin = unit(c(0,0,2,0),"cm") # remove plot margins
     )
```

```{r epicurve_facet, fig.height=8, fig.width=6.5, fig.cap='Weekly Case Count by Region, Guinee'}

# epicurve by week, Region (2014 only) 
load('regionsSL.rda')
Region = regions %>%  # table includes Region, DistrictRes, SCRes
  select(-SCRes) %>% # remove SCRes
  unique()  # collapse to unique values of Region and DistrictRes


# facet by region; fill by prefecture
cases %>% 
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter( year(dateOnset) == 2014 ) %>%
  arrange(dateOnset) %>%
  mutate( interval = week(dateOnset),
          period = as.Date(floor_date(dateOnset, "week")) ) %>%  # set time period
  group_by( Region, DistrictRes, period ) %>%
  summarize( freq = n() ) %>%  
  ggplot( aes(x = period , y = freq, fill = DistrictRes) ) +
  geom_bar( stat = 'identity', color = 'white') +
  scale_x_date('Week', labels = date_format("%b-%d"),
               breaks = date_breaks("week")) +
  ylab('Cases') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1) ) +
#   scale_fill_manual(values = c('red', 'blue', 'yellow')) +
  facet_grid( Region ~ .)


ggsave('epicurve3_SL.png', width=8, height=6 )

```

```{r tab_district, results='asis'}
p.district = cases %>%  # tabulate margin totals by DistrictRes
  filter(!is.na(period)) %>%
  group_by(DistrictRes) %>% 
  summarize(Total = n()) 

p = cases %>%  # tabulate totals by location
  filter(!is.na(period)) %>%
  group_by( period, DistrictRes ) %>% 
  summarize(Total = n()) %>%
  as.data.frame() %>%  # spread works more reliably on data.frame
  spread(period, Total)  %>%
  merge(p.district) 

p = p %>%
    arrange(DistrictRes )  # sorted by number in must recent week

# tabulate margin totals over status
TOTALS = cbind(DistrictRes = 'TOTAL',
                      data.frame( t(sapply(p[,2:ncol(p)], sum, na.rm=TRUE) ) )
                 )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(p)  

p = rbind(p, TOTALS) # join TOTALS to t

p[is.na(p)] <- 0  # convert NA to 0

kable(p)  # print table

```

```{r prefets, fig.cap="Cas d'Ebola pendant le dernier 21 jours, Guinee", fig.height=8, fig.width=6.5, eval=FALSE}

# add names of all prefets to case count table
p.name = data.frame(DistrictRes = unique(regions$DistrictRes))
p = merge(p.name, p, all.x=TRUE)
p[is.na(p)] <- 0  # convert NA to 0

# classify values
p$Cases1 =  cut( p[, 'Last 7 days'], 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)

# classify values over past 3 weeks
p$weeks3 = 
    p[, 'Last 7 days'] + p[, 'Previous 8-14 days'] + p[, 'Previous 15-21 days']
      
p$Cases3 =  cut( p[, 'weeks3'], 
                      breaks = c(0,1,10, 20, 100, Inf), 
                      label = c('0','1-9','10-19','20-100', '100 +'),
                      right=FALSE, include.lowest=TRUE)

# Join data with map
p$NAME_2 = p$DistrictRes
map.data.p = merge(SLE2, p, all.x=TRUE, 
                 by.y='NAME_2', by.x = 'NAME_2') %>%
  merge(SLE2.c, all.x = TRUE) %>%
  arrange(order) 

map.data.p.c = merge(SLE2.c, p, all.x=TRUE, by.y='NAME_2', by.x = 'Name') 

# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(5, color.brewer)

# past 3 weeks
g = ggplot( ) + theme_map +
  geom_polygon(data= map.data.p,
               aes(x=long, y=lat, group=group, 
                   fill=Cases3), 
               color="grey50", alpha=1) + 
  geom_text(data= SLE2.c, aes(label = Name, x=V1, y=V2), size=3) +
  scale_fill_manual('Cases', values = colors, drop = FALSE)
g
ggsave('PrefetMap3_SL.png')

```

```{r prefets_dot, fig.cap="Cas d'Ebola pendant le dernier 21 jours, Guinee", fig.height=8, fig.width=6.5}

# classify values
p$Cases1 =  cut( p[, 'Last 7 days'], 
                      breaks = c(0, 1, 5, 10, 20, 50, Inf), 
                      label = c('0','1-4','5-9','10-19', '20-49', '50+'),
                      right=FALSE, include.lowest=TRUE)

# classify values over past 3 weeks
p$weeks23 = 
    p[, 'Previous 8-14 days'] + p[, 'Previous 15-21 days']
      
p$Cases23 =  cut( p[, 'weeks23'], 
                      breaks = c(0,1,10, 20, 50, 100, Inf), 
                      label = c('0','1-9','10-19','20-50','50-100', '100 +'),
                      right=FALSE, include.lowest=TRUE)

# Join data with map
p$NAME_2 = p$DistrictRes
map.data.p = merge(SLE2, p, all.x=TRUE, 
                 by.y='NAME_2', by.x = 'NAME_2') %>%
  merge(SLE2.c, all.x = TRUE) %>%
  arrange(order) 

map.data.p.c = merge(SLE2.c, p, all.x=TRUE, by.y='NAME_2', by.x = 'Name') 

color.brewer ="YlOrBr" # "Purples" # 
colors = brewer.pal(6, color.brewer)  # fill colors

g = ggplot( ) + theme_map +
  geom_polygon(data= map.data.p,
               aes(x=long, y=lat, group=group, 
                   fill=Cases23), 
               color="grey50", alpha=1) + 
  geom_text(data= SLE2.c, aes(label = Name, x=V1, y=V2), size=4, hjust = .5, vjust = 1.5) +
  scale_fill_manual('8-21 days prior', values = colors, drop = FALSE) +
  geom_point(data= map.data.p.c, aes(x=V1, y=V2, size = Cases1, alpha = Cases1), 
             color = "#980043") + # maroon ""
  scale_size_discrete("1-7 days prior ") +
  scale_alpha_discrete("1-7 days prior ")
g
ggsave('PrefetMap4_SL.png', width=6, height=5, units='in')
```

# Cases by Sous Prefecture

```{r tab_sous_district, results='asis'}

ps.SCRes = cases %>%  # tabulate margin totals by DistrictRes
  filter(!is.na(period)) %>%
  group_by(DistrictRes, SCRes) %>% 
  summarize(Total = n()) 


ps = cases %>%  # tabulate totals by location
  filter(!is.na(period)) %>%
  group_by( period, DistrictRes, SCRes ) %>% 
  summarise(Total = n()) %>%
  as.data.frame() %>%
  spread(period, Total)  %>%
  merge(ps.SCRes) 


# tabulate margin totals over status
TOTALS = cbind(DistrictRes = 'TOTAL', SCRes = '',
                      data.frame( t(sapply(ps[,3:ncol(ps)], sum, na.rm=TRUE) ) )
               )

# before adding to bottom of table, t, need to be sure has same colnames
colnames(TOTALS) = colnames(ps)  

ps = rbind(ps, TOTALS) # join TOTALS to t

missing.SCRes = ps %>% filter(is.na(SCRes))

ps[is.na(ps)] <- 0  # convert NA to 0

ps[ps$SCRes == '0', 'SCRes'] = 'not_reported'

# kable(ps)  # print table
```

```{r}
# write number of cases with SC not reported.
cat("Sous-prefecture data is missing for", 
    sum(is.na(cases$SCRes)), 
    'records overall, and was not reported for', 
    nrow(missing.SCRes),
    'of', 
    ps %>% filter(DistrictRes == 'TOTAL') %>% 
                  select(3:5) %>% sum(na.rm = TRUE),
    'during past 3 weeks. \n')
```

```{r sous_prefets_3wks, fig.cap="Cas d'Ebola pendant le dernier 21 jours, Guinee", fig.height=8, fig.width=6.5, eval=FALSE}

# add names of all prefets to case count table
p.name = data.frame(SCRes = unique(SLE3$NAME_3))
ps = merge(p.name, ps, all.x=TRUE)
ps[is.na(ps)] <- 0  # convert NA to 0

# classify values over past 3 weeks
ps$weeks3 = 
    ps[, 'Last 7 days'] + ps[, 'Previous 8-14 days'] + ps[, 'Previous 15-21 days']
  

# classify values
ps$Cases3 =  cut( ps$weeks3, 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)


# Join data with map
ps$NAME_2 = ps$DistrictRes
ps$NAME_3 = ps$SCRes
map.data.ps = merge(SLE3, ps, all.x=TRUE, 
                 by.y= c('NAME_2', 'NAME_3'), 
                 by.x = c('NAME_2', 'NAME_3')
                 )%>%
  arrange(order)


# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(4, color.brewer)

# last 3 weeks
g = ggplot( ) +  theme_map # + theme_bare 
g = g + theme( axis.text.y = element_blank() )                    
g = g + geom_polygon(data= map.data.ps, 
                     aes(x=long, y=lat, group=group, fill=Cases3),
                     color="grey90", alpha=1)
g = g + geom_path(data= map.data.p, 
                     aes(x=long, y=lat, group=group),
                  color="grey20", alpha=1) 
g = g + geom_text(data= SLE2.c, aes(label = Name, x=V1, y=V2), size=3)
g = g + scale_fill_manual('Cases', values = colors, drop = FALSE)
g

ggsave('sousPrefetMap3_SL.png')
```

```{r sous_prefets_total, fig.cap="Cas d'Ebola depuis le debut d'epidemie, Guinee", fig.height=8, fig.width=6.5, eval=FALSE}
# add names of all prefets to case count table
p.name = data.frame(SCRes = unique(SLE3$NAME_3))
ps = merge(p.name, ps, all.x=TRUE)
ps[is.na(ps)] <- 0  # convert NA to 0

# classify values
ps$CasesTotal =  cut( ps[, 'Total'], 
                      breaks = c(0,1,10,20, Inf), 
                      label = c('0','1-9','10-19','20 +'),
                      right=FALSE, include.lowest=TRUE)


# Join data with map
ps$NAME_2 = ps$DistrictRes
ps$NAME_3 = ps$SCRes
map.data.ps = merge(SLE3, ps, all.x=TRUE, 
                 by.y= c('NAME_2', 'NAME_3'), 
                 by.x = c('NAME_2', 'NAME_3')
                 )%>%
  arrange(order)



# define colors
color.brewer = "PuRd" 
# "YlGnBu" , 'PuBu'  'Greys' (http://colorbrewer2.org/)
# if categorization is 'diverging'use  color.brewer = 'PiYG'
colors = brewer.pal(4, color.brewer)

g = ggplot( ) +  theme_map # + theme_bare 
g = g + theme( axis.text.y = element_blank() )                    
g = g + geom_polygon(data= map.data.ps, 
                     aes(x=long, y=lat, group=group, fill=CasesTotal),
                     color="grey90", alpha=1)
g = g + geom_path(data= map.data.p, 
                     aes(x=long, y=lat, group=group),
                  color="grey20", alpha=1) 
g = g + geom_text(data= SLE2.c, aes(label = Name, x=V1, y=V2), size=3)
g = g + scale_fill_manual('Cases', values = colors, drop = FALSE)
g

ggsave('sousPrefetMap_all_SL.png')
```

# Personnel de sante infecté et décedé par suite d'Ebola

```{r hcw_data}

# chart of hcw infection by type of work

h = cases %>% 
  filter( HCW == TRUE) %>% # limit to health care worker (HCW)
  mutate(interval = week(dateOnset), # set time period
         hcw.type = # group types hcw
           ifelse( grepl('LAB', toupper(HCWposition)), 'Lab',
              ifelse( grepl('M.D', toupper(HCWposition)),
                           'Medecin', 
                ifelse( grepl('INF', toupper(HCWposition))
                                   , 'Infirmiere', 
                                  'Other'))),
         frequency = 1
         ) %>%  
  filter( year(dateOnset) == 2014 ) %>%
  select( DistrictRes, SCRes, interval, hcw.type, HCWposition, StatusAsOfCurrentDate, frequency) %>%
  merge(regions, all.x = TRUE) %>%
  arrange(interval, hcw.type ) 
```

Since the onset of the epidemic, there have been 
`r nrow(h) ` cases among health care workers.

```{r chart_hcw_data, fig.height=8, fig.width=6.5}

# point shapes in each box
h.2 = h %>%
  select(interval, DistrictRes, hcw.type, StatusAsOfCurrentDate, frequency) %>% 
  arrange(interval, hcw.type, StatusAsOfCurrentDate) %>%
  group_by(interval, DistrictRes) %>%
  mutate( 
    cumsum = cumsum( frequency)  ,
    midpoint = cumsum - .5
    )

h.2 %>%
  ggplot( aes(x = interval ) ) +
  geom_bar( aes(y = frequency, fill = hcw.type), 
            stat = 'identity', color = 'white') +
  geom_point(data = h.2, 
             aes(y = midpoint, ymax = midpoint, 
                 shape = StatusAsOfCurrentDate, 
                 alpha = StatusAsOfCurrentDate)
             ) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = 'top' ) +
  scale_x_discrete('Week', limits = c(min(h$interval) - 1, max(h$interval) + 1) ) +
  scale_fill_discrete('Position') +
  scale_shape_manual('Status', values = c(25,19,19)) +
  scale_alpha_manual('Status', values = c(.5, 0,1)) +
  facet_grid( DistrictRes ~ .) 

ggsave('epicurve_HCW_SL.png', width=8, height=6 )

```

Of these, 
`r nrow(h %>% filter(StatusAsOfCurrentDate == 'Décédé'))` died.

```{r hcw_risk}
t = with(h, table(hcw.type, StatusAsOfCurrentDate))
prop.table(t, 1)

t = with(h %>% filter(hcw.type == 'Other'), table(HCWposition, StatusAsOfCurrentDate))
t.p = prop.table(t, 1)
kable(t)
```


For the sake of visualiztion, data avaialable on the type health work was classified as Infirmiere, Lab, Medecin, or other.  

The following table lists the position descriptions available for type of health care worker.  

`r kable(h %>% count(hcw.type))`

The free text given for each type is shown below.  


Infirmiere: 
```{r}
I = h  %>%  
  filter(hcw.type %in% 'Infirmiere')

paste( unique(I$HCWposition), collapse = ', ') 
```

Lab: 
```{r}
L = h  %>%  
  filter( hcw.type %in% 'Lab')

paste( unique(L$HCWposition), collapse = ', ') 
```

Medecin: 
```{r}
M= h  %>%  
  filter(hcw.type %in% 'Medecin')

paste( unique(M$HCWposition), collapse = ', ') 
```

Other: 
```{r}
O = h  %>%  
  filter( hcw.type %in% 'Other')

paste( unique(O$HCWposition), collapse = ', ') 
```


# Days to Isolation

The days to isolation was calculated as the number of days from illness onset to hospital admission. To assess changes, we plot the days to isolation by month, and by month and region.

```{r isolation_histo, fig.cap = 'Histogram of days to isolation'}

# Time from onset to isolation
d = cases %>% 
  select( DateOnset, DateIsolationCurrent, StatusAsOfCurrentDate, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ) ) 

d %>% ggplot(aes( daysToIso)) + geom_bar()
```

+ Because of data entry errors, some values for days to hospitalization do not make sense. 

+ For this analysis, days to isolation were restricted to 0-20 days (results: removed ~7% of reports) 

```{r isolation, fig.cap='Days to Isolation by Month of Onset, SL'}

d_filter = d %>% filter(daysToIso >= 0 & daysToIso < 21 & 
                          StatusAsOfCurrentDate %in% c('ALIVE', 'DEAD')) 

ggplot(d_filter, aes( month(DateOnset), daysToIso)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c( "green", "red") ) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA) 

# d$daysToIso_range = cut(d$daysToIso, c(-Inf, 0, 21, Inf ))
# t = table(d$daysToIso_range)
# prop.table(t)

```


```{r isolation_region, fig.height=8, fig.cap='Days to Isolation by Month of Onset, Region, SL'}

d_filter = d %>% filter(daysToIso >= 0 & daysToIso < 21 & 
                          StatusAsOfCurrentDate %in% c('ALIVE', 'DEAD')) 

ggplot(d_filter, aes( month(DateOnset), daysToIso)) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    scale_y_continuous('Days to Isolation', breaks = seq(0, 28, 2)) +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c( "green", "red") ) +
    geom_boxplot( aes(group = month(DateOnset)) , outlier.shape = NA, fill = NA) +
    facet_grid( Region ~ .)
```


# Survival as function of days to isolation

To assess the impact of contact tracing and access to care on survival, we examined association of days to isolation with survival ('StatusAsOfCurrentDate')

+ Because the data on survival is typically provided at the outset of hospitalization, the final outcome may not be updated, particularly for the most recent records. Therefore many outcomes may be misclassified as survived even thought the patient died.  Older records are more likely to be correct. 

+ For this analysis, days to isolation were restricted to 0-20 days  and did not include last month of reports (because survival data likely not updated).  

```{r isolation_survival, fig.cap='Days to Isolation by Month of Onset, SL'}

# remove last month because data not updated
d_filter2 = d_filter %>% filter(month(DateOnset) < 12)


d_filter2 %>% filter(StatusAsOfCurrentDate %in% 'DEAD' ) %>% 
  select(daysToIso_Died = daysToIso) %>% summary(daysToIso) %>% kable()


d_filter2 %>% filter(StatusAsOfCurrentDate %in% 'ALIVE' ) %>% 
  select(daysToIso_Survived = daysToIso) %>% summary(daysToIso) %>% kable()


# overall
ggplot(d_filter2, aes( StatusAsOfCurrentDate, daysToIso)) +
    scale_y_continuous('Days to Isolation', breaks = c(1:20)) +
    scale_x_discrete('Status') +
    geom_point(aes(colour = StatusAsOfCurrentDate), position = 'jitter' ) +
    scale_colour_manual('Status', values = c( "green", "red") )  +
    geom_boxplot(  outlier.shape = NA, fill = NA)

```

## Survival rate

```{r isolation_survival_rate, fig.cap='Survival rate (loess) by days to isolation, SL'}

# remove last month because data not updated
d_filter2_rate = d_filter2 %>%
  group_by(daysToIso) %>%
  summarize( survival = sum(StatusAsOfCurrentDate == 'ALIVE')/ n(),
             n = n())

# overall
ggplot(d_filter2_rate, aes( daysToIso, survival)) +
    geom_point(aes(size = n)) +
    scale_x_continuous('Days to Isolation', breaks = c(1:20)) +
    scale_y_continuous('Survival Rate', limits = c(0,1), labels = percent ) +
    geom_smooth(aes(weight = n)) 

```



```{r isolation_survival_rate_region, fig.cap='Survival rate (loess) by days to isolation, SL'}

# remove last month because data not updated
d_filter3_rate = d_filter2 %>%
  # filter( Region %in% c('Conakry', 'Nzérékoré', 'Faranah')) %>%
  group_by(daysToIso, Region) %>%
  summarize( survival = sum(StatusAsOfCurrentDate == 'ALIVE')/ n(),
             n = n())

# overall
ggplot(d_filter3_rate, aes( daysToIso, survival)) +
    geom_point(aes(size = n)) +
    scale_x_continuous('Days to Isolation', breaks = c(1:20)) +
    scale_y_continuous('Survival Rate', limits = c(0,1), labels = percent ) +
    geom_smooth(aes(weight = n)) + facet_grid( Region ~ .)

```

# Number Not Isolated

```{r noIso}

# Time from onset to isolation
data = cases %>% 
  select( DateOnset, DateIsolationCurrent, StatusReport, DistrictRes) %>%
  merge(Region, all.x = TRUE, by= 'DistrictRes') %>%
  filter(DateIsolationCurrent > DateOnset ) %>%
  mutate( daysToIso = as.numeric( as.Date(DateIsolationCurrent) - as.Date(DateOnset) ),
          StatusReport = factor(StatusReport, labels = c('Dead', 'Alive'))
          ) %>%
  filter(daysToIso < 28)
 
g = ggplot(data, aes( month(DateOnset))) +
    scale_x_continuous('Month', breaks = c(1:12)) +
    geom_bar(aes(fill = StatusReport), color = 'black', stat='bin', binwidth = 1, position = 'stack' ) +
    scale_fill_manual('Status at Report', values = c("red", "green")) 
g

kable( data %>% filter(StatusReport == 'Dead') %>% count(Region))
```

