---
title: "Guinee-VHF Variable list"
author: "CDC/OMS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes
  word_document:
  - toc: yes
  
header-includes:
- \usepackage[section]{placeins}
- \usepackage{setspace}
- \usepackage{relsize}
---

```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE }
options(tz="GMT+0")
Sys.setenv(TZ='GMT+0')

library(knitr)
opts_chunk$set( echo=FALSE, message=FALSE, comment="", 
                warnings=FALSE, results='asis',
                cache = FALSE)
library(mi)
library(Hmisc)
library(XLConnect, quietly=TRUE)
library(lubridate)
library(plyr)

library(ggplot2)
library(scales)
library(RColorBrewer)

library(tidyr)
library(dplyr, quietly=TRUE) # order is important...

library(rgeos)
```

# Cases

- Y compris les cas confirme, probable, et suspect

```{r read_cas}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.

dir = 'fieldData/'

file.name = "guinea_cases_10_Nov"

excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
case_data = readWorksheet(wb, file.name)

# csv.file = paste0(dir, file.name, '.csv')
# case_data = read.csv(csv.file, header=TRUE)
```

```{r Define_case_data}

epiCaseDef = data.frame(EpiCaseDef = c(0,1,2,3,4),
                     EpiCaseDef.label = c('Non-Cas','Confirme',
                                          'Probable','Suspect', 'Exclude'))

# if else with dates can be troublesome.  see sol at 
# http://stackoverflow.com/questions/6668963/how-to-prevent-ifelse-from-turning-date-objects-into-numeric-objects
date.ifelse <- function(cond, yes, no){ class.y <- class(yes)
                    ø              X <- ifelse(cond,yes,no)
                                  class(X) <-class.y; return(X)}


last.report.date = max(case_data$DateReport, na.rm=T) # last day with report
current.week = week(last.report.date)  

cases = case_data %>% 
  select( ID, StatusAsOfCurrentDate, 
          EpiCaseDef , StatusReport, FinalStatus,
          ThisCaseIsAlsoContact, TotalContactsListed, 
          DateReport, DateOnset,
          DateHospitalCurrentAdmit, DateIsolationCurrent,
          Age, AgeUnit, Gender, DateDeath, 
          VillageRes, ParishRes, CountryRes,
          DistrictRes, SCRes, HCW, HCWposition, HCWFacility, 
          Funeral, Animals, 
          Contact, ContactName1, ContactRelation1, ContactDateStart1, ContactDateEnd1,
          ContactName2, ContactRelation2,  ContactDateStart2, ContactDateEnd2
          ) %>% 
  filter(!EpiCaseDef==4) %>% # drop 'excluded'
  mutate( EpiCaseDef = as.integer(EpiCaseDef)) %>%
  inner_join(epiCaseDef) %>%  
  mutate( 
          dateOnset = 
            date.ifelse( is.na(DateOnset), DateReport, DateOnset )
            , # use onset date if available; DateReport otherwise
          week = week(dateOnset)  ,
          # calculate periods based on week function, starting with day before current date
          period_wk = ifelse( week == current.week, "Week", 
                          ifelse( week == ( current.week - 1 ), "Week_1",
                                 ifelse( week == ( current.week - 2 ), "Week_2",
                                        ifelse( week == ( current.week - 3 ), "Week_3",
                                        ifelse( week== ( current.week - 4 ), 
                                               "Week_4", "Previous"))))),
          period_wk = factor(period_wk, 
                          levels = c('Week','Week_1','Week_2', 'Week_3', 'Week_4', 'Previous') ),
          
          # calculate periods based on 7 day intervals, 
          # starting with day before current date
          period = date.ifelse( dateOnset %within% 
                             interval(last.report.date-edays(6),
                                      last.report.date
                                      ), "Last 7 days", 
                    date.ifelse(dateOnset %within% 
                           interval(last.report.date-edays(13), 
                                    last.report.date-edays(7)), "Previous 8-14 days",
                    date.ifelse(dateOnset %within% 
                             interval(last.report.date-edays(20), 
                                      last.report.date-edays(14)), "Previous 15-21 days", 
                           "Previous"))),
          period = factor(period, 
                          levels = c('Last 7 days','Previous 8-14 days',
                                     'Previous 15-21 days', 'Previous') ),
          onset.last21days = ifelse( dateOnset  %within% 
                                     interval(Sys.Date()-edays(1), 
                                              Sys.Date()-edays(22)), 
                                     TRUE, FALSE) ,
          CaseMonth = paste(month(dateOnset), year(dateOnset)),
          Status = EpiCaseDef.label
          ) %>%
    filter( Status %in% c('Confirme','Probable', 'Suspect'))

```

```{r missing_date, results='asis'}
#  list records missing period
missingPeriod = cases %>% filter(is.na(period)) %>% select(ID)

cat('There are', nrow(missingPeriod), 'records with missing date:\n')

cat(missingPeriod$ID, sep=', ')

```

Focus on cases since 9-1-2104.

- There were `r nrow(cases %>% filter( DateReport > ymd("2014-8-31")))` cases before September 1, and

- There have been `r nrow(cases %>% filter( DateReport <= ymd("2014-8-31")))` since Septmber 1.

## Missing data via mi package

```{r mi, fig.height=8}

cases1 = cases %>% filter( DateReport <= ymd("2014-8-31"))

cases2 = cases %>% filter( DateReport > ymd("2014-8-31"))

par(mai=c(1,2,1,0))
missing.pattern.plot( cases1 ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(cases1), ")" ),
                      main = "Cases reports before Sept 1")

missing.pattern.plot( cases2 ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(cases2), ")" ), 
                      main = "Case reports after Sept")

mis =  mi.info(cases2) 
miss = data.frame(mis$name, mis$nmis, mis$type)
colnames(miss) = c('Variable', 'Nmiss', 'Type')
kable(miss, row.names=F)

```

## Liste des variables condensées dans la table de cas

```{r describe_cases}

# Cases = cases2 %>% select(-contains('ID'), -contains('GlobalRecordId') )
d = describe(cases2)
latex(d, append=TRUE, file='')
```

# Contacts

```{r read_contacts}

# In order for this import to work, must do 2 things first:
# 1. Make sure Region settings (control panel) are set to English(UK)
# 2. open csv file in excel, then save as an excel spreadsheet.

file.name = "guinea_contacts_10_Nov"

dir = 'fieldData/'

excel.file = paste0(dir, file.name, '.xlsx') 
wb = loadWorkbook(excel.file)
contact_data = readWorksheet(wb, file.name)

contacts = contact_data %>%
  mutate(ContactMonth = paste(month(DateLastContact), year(DateLastContact)) )

save(contacts, file = "contacts.rda")
```

```{r mi_contacts, fig.height=8}
par(mai=c(1,2,1,0))
missing.pattern.plot( contacts ,clustered=FALSE,
                      xlab= paste0("observations (n=", nrow(contact_data), ")" ),
                      main = "contacts")

miss = data.frame(mis$name, mis$nmis, mis$type)
colnames(miss) = c('Variable', 'Nmiss', 'Type')
kable(miss, row.names=F)

```


```{r describe_contacts}

# contacts = contact_data %>% select(-ID)
d = describe(contacts)
latex(d, append=TRUE, file='')
```


